---
title: "Eyetracking QC Visualizations"
output: html_notebook
---

This notebook generates quality control visualizations from the QC report CSV.

# Setup

```{r setup, message=FALSE}
library(tidyverse)
library(here)
library(forcats)


# Specify which QC report to load
qc_filename <- "qc_report_20260511.csv"  # <-- CHANGE THIS TO YOUR FILE

# Load the data
qc_data <- read_csv(here("reports/", qc_filename)) |> filter(participant_id != "MW000")

# Quick look at the data
glimpse(qc_data)
```

# Plot 1a: Validity by Participant

One bar per participant showing mean % valid (averaged across both eyes and all trials), with error bars reflecting variance across trials.

```{r validity-by-participant, fig.width=12, fig.height=5}
participant_summary <- qc_data %>%
  mutate(participant_type = case_when(
    startsWith(participant_id, "MWK") ~ "kids (5-7 yos)",
    startsWith(participant_id, "MW") ~ "adults",
    startsWith(participant_id, "HMET") ~ "infants (12-24 mos)"
  )) %>%
  group_by(participant_id, participant_type) %>%
  summarise(
    mean_valid = mean(mean_percent_valid),
    sd_valid = sd(mean_percent_valid),
    n_trials = n(),
  )

ggplot(participant_summary, aes(x = participant_id, y = mean_valid)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_errorbar(aes(ymin = mean_valid - sd_valid, 
                    ymax = mean_valid + sd_valid), 
                width = 0.2) +
  labs(
    title = "Data Validity by Participant",
    x = "Participant",
    y = "Mean % Valid (both eyes)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
# Plot 1b: Validity by Participant Group

```{r validity-by-participant-type, fig.width=12, fig.height=8}
summarized_data <- function(data, x_var, y_var, group_var) {
  return(data %>%
           group_by_at(c(x_var, group_var)) %>%
           summarise(mean_value = mean(.data[[y_var]], na.rm = TRUE),
                     sd_value = sd(.data[[y_var]], na.rm = TRUE),
                     n = n(),
                     se = sd_value / sqrt(n()),
                     ci_lower = mean_value - qt(1 - (0.05 / 2), n - 1) * se,
                     ci_upper = mean_value + qt(1 - (0.05 / 2), n - 1) * se,
                     .groups = 'drop')
  )
}

participant_type_summary <- summarized_data(participant_summary |> filter(mean_valid > 50), "participant_type", "mean_valid", "participant_type") 

participant_type_summary$participant_type <-
  fct_relevel(participant_type_summary$participant_type, "adults", after = Inf)


participant_type_summary_plot <- ggplot(participant_type_summary, aes(x=participant_type, y=mean_value)) +
  geom_point(size=4) +
  geom_errorbar(aes(ymin=ci_lower, ymax=ci_upper), width=0) +
  geom_jitter(data=participant_summary, aes(x=participant_type, y=mean_valid), alpha=0.5, width=0.1) +
  ylab("Mean valid % (both eyes)") +
  geom_hline(yintercept=50, linetype="dotted") +
  xlab("Participant type")

ggsave(here("figures/valid_data_per_group.png"), participant_type_summary_plot)
```

# Plot 2a: Validity by Trial Name (stimulus)

Shows validity trends by stimulus, with each participant as a separate line.

```{r validity-by-trial-name, fig.width=10, fig.height=6}
ggplot(qc_data, aes(x = trial_name, y = mean_percent_valid, 
                    color = participant_id, group = participant_id)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Data Validity by Stimulus",
    x = "Trial Name",
    y = "% Valid (both eyes)",
    color = "Participant"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
```{r}
mw031_data <- qc_data |> filter(participant_id == 'MW031') |>
  pivot_longer(cols=c(left_percent_valid, right_percent_valid), names_to="eye", values_to="percent_valid")
ggplot(mw031_data, aes(x = trial_name, y = percent_valid, 
                    color = eye, group = eye)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Data Validity by Stimulus",
    x = "Trial Name",
    y = "% Valid (left/right)",
    color = "Eye"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Plot 2b: Validity by Trial Index (over time)

Shows validity trends over the course of the experiment, with each participant as a separate line.

```{r validity-by-trial-index, fig.width=10, fig.height=6}
ggplot(qc_data, aes(x = trial_index, y = mean_percent_valid, 
                    color = participant_id, group = participant_id)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Data Validity Over Time",
    x = "Trial Index",
    y = "% Valid (both eyes)",
    color = "Participant"
  ) +
  scale_x_continuous(breaks = seq(0, max(qc_data$trial_index), 1)) +
  theme_minimal()
```
```{r}
qc_data_trial_mean <- qc_data |> 
  group_by(trial_index) |>
  summarize(
    group_mean_valid = mean(mean_percent_valid),
    ci = qt(0.975, df = n() - 1) * sd(mean_percent_valid) / sqrt(n())
  )
n_participant <- length(unique(qc_data$participant_id))

ggplot(qc_data, aes(x = trial_index, y = mean_percent_valid, 
                    color = as.numeric(str_extract(participant_id, "\\d+$")), group = participant_id)) +
  geom_line(alpha = 0.4) +
  geom_line(data = qc_data_trial_mean, color = "black", aes(y = group_mean_valid, group = NA)) +
  geom_ribbon(data = qc_data_trial_mean, 
              aes(y = group_mean_valid, 
                  ymin = group_mean_valid - ci, 
                  ymax = group_mean_valid + ci,
                  group = NA),
              fill = "black", alpha = 0.2, color = NA) +
  labs(
    title = "Data Validity Over Time",
    x = "Trial Index",
    y = "% Valid (both eyes)",
    color = "Participant"
  ) +
  scale_color_viridis_c(
    name = "Participant",
    breaks = seq(1, n_participant, by = 4),
    labels = seq(1, n_participant, by = 4),
    limits = c(1, n_participant)
  ) +
  scale_x_continuous(breaks = seq(0, max(qc_data$trial_index), 1)) +
  theme_minimal() +
  guides(color = guide_colorbar(barwidth = 4, barheight = 5, title.position = "top"))

```

# Plot 3: Calibration Drift Over Blocks

Shows validation accuracy (in degrees) across blocks, with one line per participant. Higher values = worse calibration.

```{r calibration-drift, fig.width=8, fig.height=5}
# Get one row per participant per block (they're repeated across trials)
block_calibration <- qc_data %>%
  select(participant_id, block_index, val_mean_deg) %>%
  distinct() 

ggplot(block_calibration, aes(x = block_index, y = val_mean_deg, 
                               color = participant_id, group = participant_id)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Calibration Drift Across Blocks",
    x = "Block Index",
    y = "Validation Accuracy (degrees)",
    color = "Participant"
  ) +
  scale_x_continuous(breaks = 0:3) +
  theme_minimal()
```

# Plot 4: Calibration vs Validity

Scatter plot showing whether worse calibration (higher degrees) is associated with more data loss (lower % valid). Each dot is one trial, colored by participant.

```{r calibration-vs-validity, fig.width=8, fig.height=6}
ggplot(qc_data, aes(x = val_mean_deg, y = mean_percent_valid, 
                    color = participant_id)) +
  geom_point(size = 2, alpha = 0.7) +
  labs(
    title = "Calibration Accuracy vs Data Validity",
    x = "Validation Accuracy (degrees, lower = better)",
    y = "% Valid (both eyes)",
    color = "Participant"
  ) +
  theme_minimal()
```

# Summary Table

```{r summary-table}
summary_data <- qc_data %>%
  group_by(participant_id) %>%
  summarise(
    n_trials = n(),
    mean_validity = round(mean(mean_percent_valid), 1),
    min_validity = round(min(mean_percent_valid), 1),
    mean_calibration_deg = round(mean(val_mean_deg), 2),
    max_calibration_deg = round(max(val_mean_deg), 2)
  ) 
  
summary_data |>
  knitr::kable()

write.csv(summary_data, here("reports/summary.csv"))
```
