---
title: "Inter-Subject Correlation (ISC) — Movie Watching Eye-Tracking"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    highlight-style: github
---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(glue)
library(patchwork)
library(zoo)
library(here)
library(ggthemes)

# ── ISC source ────────────────────────────────────────────────────────────────
# "gaze": calc_isc_rawgaze.py output (raw continuous gaze) -- needs the QC filter in
#         Part 1 below, since validity/trial-count screening hasn't happened
#         upstream of that script.
# "fixations": calc_isc_fixations.py output (I2MC fixation-position ISC) -- Part 1
#         is skipped, since trial-level exclusion (calibration accuracy,
#         fraction of invalid samples) already happened during fixation
#         detection itself (i2mc_fixations.py's max_calibration_deg /
#         max_invalid_frac), before a pid/video pair could ever reach
#         isc_fixation_pairwise.csv.
isc_source <- "fixations"   # "gaze" or "fixations"

# ── Paths ─────────────────────────────────────────────────────────────────────
SERVER_PATH <- "/Volumes/vislearnlab/experiments/movie-watching"
data_dir         <- file.path(SERVER_PATH, "data/raw/")
isc_pairwise_csv <- file.path(
  SERVER_PATH,
  if (isc_source == "fixations") "data/results/isc/isc_fixation_pairwise.csv"
  else                           "data/results/isc/isc_pairwise.csv"
)
summary_csv      <- here("reports/summary.csv")

# ── QC thresholds 
min_valid_pct <- 50   # mean_validity % across trials
min_trials    <- 6    # minimum usable trials (half of total trials)

# ── Example time-series participants 
# Two adults and one child used for the 10-second gaze trace plot.
# Change these to any valid participant IDs in your dataset.
ex_adult_1 <- "MW046"
ex_adult_2 <- "MW028"
ex_child   <- "MWK012"

example_video <- "pixar_birds"   # video shown in the time-series panel
```

---

# Part 1: QC Filtering

`isc_source == "gaze"`: we use the pre-computed `summary.csv` to identify
which participants have sufficient data quality, then filter
`isc_pairwise.csv` to only those pairs where **both** participants pass QC.

`isc_source == "fixations"`: skipped -- `isc_fixation_pairwise.csv` only
ever contains pid/video pairs that already survived per-trial exclusion
(calibration accuracy, fraction of invalid samples) during fixation
detection, so there's nothing left to filter here.

```{r qc}
if (isc_source == "gaze") {
  summ <- read_csv(summary_csv, show_col_types = FALSE) |>
    select(participant_id, n_trials, mean_validity)

  passed_qc <- summ |>
    filter(mean_validity >= min_valid_pct, n_trials >= min_trials) |>
    filter(!startsWith(participant_id, "HMET")) |> #excluding infant participants for now
    pull(participant_id)

  cat(glue("{length(passed_qc)} / {nrow(summ)} participants pass QC ",
           "(≥{min_valid_pct}% valid, ≥{min_trials} trials)\n\n"))

  summ |>
    mutate(included = participant_id %in% passed_qc) |>
    arrange(desc(included), desc(mean_validity))
} else {
  cat(glue("Skipping QC filtering -- isc_source = \"fixations\": trial-level ",
           "exclusion already happened upstream during fixation detection.\n"))
}
```

---

# Part 2: Load & Filter Pairwise ISC

```{r load-isc}
isc_raw <- read_csv(isc_pairwise_csv, show_col_types = FALSE)

# Keep only pairs where both participants passed QC (isc_source == "gaze"
# only -- for "fixations" every row already qualifies, see Part 1)
isc_qc <- if (isc_source == "gaze") {
  isc_raw |> filter(pid_a %in% passed_qc, pid_b %in% passed_qc)
} else {
  isc_raw
}

isc <- isc_qc |>
  # Use isc_xy (average of horizontal + vertical) as the primary ISC measure
  rename(isc = isc_xy) |>
  # Simplify comparison labels for plotting
  mutate(
    pair_type = case_when(
      comparison == "adults-adults"   ~ "Adult–Adult",
      comparison == "kids-kids"       ~ "Child–Child",
      comparison == "adults-kids"     ~ "Child–Adult",
      comparison == "infants-infants" ~ "Infant–Infant",
      comparison == "adults-infants"  ~ "Adult–Infant",
      comparison == "infants-kids"    ~ "Infant–Child",
      TRUE                             ~ comparison
    ),
    pair_type = factor(pair_type, levels = c(
      "Infant–Infant", "Infant–Child", "Child–Child",
      "Child–Adult", "Adult–Infant", "Adult–Adult"
    ))
  )

cat(glue("Retained {nrow(isc)} pair × video rows ",
         "({n_distinct(c(isc$pid_a, isc$pid_b))} unique participants)\n"))
isc |> count(pair_type)
```

---

# Part 3: Example Time-Series (`r example_video`, first 10 s)

We load the raw gaze files for the three example participants directly.

```{r ts-load-helpers}
# ── File discovery: participant ID is the prefix of the filename ──────────────
find_gaze_path <- function(pid) {
  hits <- list.files(data_dir, pattern = glue("^{pid}_.*\\.csv$"),
                     full.names = TRUE, recursive = TRUE) |>
    keep(~ !str_detect(basename(.x), "_trial_order|_validation"))
  if (length(hits) == 0) { warning(glue("No gaze file found for {pid}")); return(NULL) }
  hits[which.min(nchar(basename(hits)))]
}

find_trial_order_path <- function(pid) {
  hits <- list.files(data_dir, pattern = glue("^{pid}_.*_trial_order\\.csv$"),
                     full.names = TRUE, recursive = TRUE)
  if (length(hits) == 0) { warning(glue("No trial_order found for {pid}")); return(NULL) }
  hits[1]
}

load_gaze_pid <- function(pid) {
  path <- find_gaze_path(pid)
  if (is.null(path)) return(NULL)
  read_csv(path, show_col_types = FALSE) |> mutate(participant = pid)
}

# ── Load the three example participants ───────────────────────────────────────
example_pids <- c(ex_adult_1, ex_adult_2, ex_child)

gaze_ex <- map(example_pids, load_gaze_pid) |>
  compact() |>
  list_rbind()

# Use one participant's trial order for the video -> trial_index lookup
to_path <- find_trial_order_path(ex_adult_1)
trial_order_ex <- read_csv(to_path, show_col_types = FALSE)

video_key <- trial_order_ex |>
  select(trial_index = total_trial_index, video_name, block_id)

cat("Loaded gaze for:", paste(unique(gaze_ex$participant), collapse = ", "), "\n")
```

```{r ts-parse}
# Label rows with trial / video (same logic as before)
gaze_labeled_ex <- gaze_ex |>
  group_by(participant) |>
  mutate(
    video_name = str_extract(events, "(?<=\\|Video_)[^|]+"),
    video_name = zoo::na.locf(video_name, na.rm = FALSE)
  ) |>
  ungroup() |>
  left_join(video_key, by = "video_name") |>
  filter(!is.na(trial_time), !is.na(trial_index))

# Resample to 8 ms grid (4 ms == 250 Hz)
gaze_resampled_ex <- gaze_labeled_ex |>
  mutate(time_bin = round(trial_time / 8) * 8) |>
  group_by(participant, trial_index, time_bin) |>
  summarise(gaze_x = mean(gaze_x, na.rm = TRUE),
            gaze_y = mean(gaze_y, na.rm = TRUE),
            .groups = "drop")
```

```{r ts-plot-helper}
plot_isc_timeseries <- function(p1, p2, label1, label2, panel_title,
                                video = example_video, max_sec = 10) {
  t_idx <- video_key |> filter(video_name == video) |> pull(trial_index)
  if (length(t_idx) == 0) {
    message("Video '", video, "' not found. Available: ",
            paste(sort(unique(video_key$video_name)), collapse = ", "))
    return(NULL)
  }
  t_idx <- t_idx[1]

  ts_data <- gaze_resampled_ex |>
    filter(participant %in% c(p1, p2),
           trial_index == t_idx,
           time_bin <= max_sec * 1000) |>
    group_by(participant) |>
    mutate(gaze_x_norm = (gaze_x - min(gaze_x, na.rm = TRUE)) /
             (max(gaze_x, na.rm = TRUE) - min(gaze_x, na.rm = TRUE))) |>
    ungroup() |>
    mutate(participant = factor(participant, levels = c(p1, p2)))

  # Pull ISC value from the pre-computed pairwise table
  isc_val <- isc |>
    filter(
      (pid_a == p1 & pid_b == p2) | (pid_a == p2 & pid_b == p1),
      video_name == video
    ) |>
    pull(isc_x) |> first() |> round(2)
  isc_label <- if (is.na(isc_val)) "ISC = NA" else glue("horizontal ISC = {isc_val}")

  ggplot(ts_data, aes(x = time_bin / 1000, y = gaze_x_norm,
                      colour = participant, linewidth = participant)) +
    geom_line(na.rm = TRUE) +
    scale_colour_manual(
      values = setNames(c("grey60", "black"), c(p1, p2)),
      labels = setNames(c(label1, label2), c(p1, p2))
    ) +
    scale_linewidth_manual(
      values = setNames(c(0.7, 1.2), c(p1, p2)),
      labels = setNames(c(label1, label2), c(p1, p2))
    ) +
    annotate("text", x = max_sec * 0.97, y = 0.05,
             label = isc_label, hjust = 1, size = 3.5) +
    scale_x_continuous(breaks = seq(0, max_sec, 2)) +
    scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.05, 1.05)) +
    labs(title = panel_title, x = "Time (s)",
         y = "Horizontal gaze position",
         colour = NULL, linewidth = NULL) +
    theme_classic(base_size = 12) +
    theme(legend.position = c(0.08, 0.25))
}
```

```{r ts-plots, fig.width=8, fig.height=7, warning=FALSE}
p_aa <- plot_isc_timeseries(
  ex_adult_1, ex_adult_2, "A1", "A2",
  panel_title = glue("(a) Adult–Adult ISC ({example_video})")
)
p_ca <- plot_isc_timeseries(
  ex_adult_1, ex_child, "A1", "C1",
  panel_title = glue("(b) Child–Adult ISC ({example_video})")
)
p_aa / p_ca
ggsave(here("figures/isc_correlation.png"))
```

::: {.callout-tip}
**What to notice**

- In the adult–adult panel the traces closely follow each other — both adults tend to look at the same screen region at the same time.
- In the child–adult panel the traces diverge more, reflected in a lower ISC value.
:::

---

# Part 4: ISCw and ISCa

Following Franchak et al. (2016):

- **ISCw** — each observer's mean ISC with all *same-age* partners, averaged across videos.
- **ISCa** — each child's mean ISC with all *adult* partners, averaged across videos.

```{r isc-averages}
# Reshape to one row per focal observer × partner × video
isc_long <- isc |>
  mutate(
    partner_a = pid_b,
    partner_b = pid_a
  ) |>
  pivot_longer(
    cols = c(pid_a, pid_b),
    names_to = "role",
    values_to = "focal"
  ) |>
  mutate(
    partner = if_else(role == "pid_a", partner_a, partner_b),
    focal_group = if_else(role == "pid_a", group_a, group_b),
    partner_group = if_else(role == "pid_a", group_b, group_a)
  ) |>
    select(
    focal,
    partner,
    video_name,
    isc,
    pair_type,
    focal_group,
    partner_group
  )

# ISCw: within-age, averaged over partners then over videos
iscw_by_video <- isc_long |>
  filter(focal_group == partner_group) |>
  group_by(focal, focal_group, video_name) |>
  summarise(ISCw = mean(isc, na.rm = TRUE), .groups = "drop")

iscw_by_subj <- iscw_by_video |>
  group_by(focal, focal_group) |>
  summarise(ISCw = mean(ISCw, na.rm = TRUE), .groups = "drop")

# ISCa: each non-adult observer averaged with all adults
isca_by_video <- isc_long |>
  filter(focal_group != "adults", partner_group == "adults") |>
  group_by(focal, focal_group, video_name) |>
  summarise(ISCa = mean(isc, na.rm = TRUE), .groups = "drop")

isca_by_subj <- isca_by_video |>
  group_by(focal, focal_group) |>
  summarise(ISCa = mean(ISCa, na.rm = TRUE), .groups = "drop")
```

---

# Part 5: Summary Plots

Two versions of a Franchak et al.–style dot plot. Mean ± 1 SE shown as a
filled point + error bars; grey dots are individual observations.

**Version A** — one dot per *subject* (ISC averaged over videos per subject)  
**Version B** — one dot per *video* (ISC averaged over subjects per video)

```{r build-plot-data}
# We plot only the three pairings that mirror Franchak Fig. 2:
# kids-kids, adults-kids, adults-adults
# (infants shown separately if present)

plot_comparisons <- c("Child–Child", "Child–Adult", "Adult–Adult")

# ── Version A: per subject ────────────────────────────────────────────────────
subj_avg <- bind_rows(
  # Adult–Adult: each adult's mean ISC with all other adults
  isc_long |>
    filter(focal_group == "adults", partner_group == "adults") |>
    group_by(focal) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Adult–Adult"),

  # Child–Child
  isc_long |>
    filter(focal_group == "kids", partner_group == "kids") |>
    group_by(focal) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Child–Child"),

  # Child–Adult (from child's perspective)
  isca_by_subj |>
    filter(focal_group == "kids") |>
    rename(isc_mean = ISCa) |>
    select(focal, isc_mean) |>
    mutate(pair_type = "Child–Adult")
) |>
  mutate(pair_type = factor(pair_type, levels = plot_comparisons))

# ── Version B: per video ──────────────────────────────────────────────────────
video_avg <- isc |>
  filter(pair_type %in% plot_comparisons) |>
  group_by(pair_type, video_name) |>
  summarise(
    isc_mean = mean(isc, na.rm = TRUE),
    se_isc   = sd(isc, na.rm = TRUE) / sqrt(sum(!is.na(isc))),
    .groups  = "drop"
  ) |>
  mutate(pair_type = factor(pair_type, levels = plot_comparisons))

# ── Summary stats ─────────────────────────────────────────────────────────────
se_stats <- function(df) {
  df |>
    group_by(pair_type) |>
    summarise(
      mean_isc = mean(isc_mean, na.rm = TRUE),
      se_isc   = sd(isc_mean, na.rm = TRUE) / sqrt(sum(!is.na(isc_mean))),
      .groups  = "drop"
    )
}

subj_stats  <- se_stats(subj_avg)
video_stats <- se_stats(video_avg)
```

```{r plot-theme}
isc_theme <- theme_few(base_size = 13) +
  theme(
    axis.text.x        = element_text(size = 11),
    legend.position    = "none",
    panel.grid.major.y = element_line(colour = "grey92")
  )

dot_plot <- function(stats_df, jitter_df, plot_title) {
  ggplot(stats_df, aes(x = pair_type, y = mean_isc)) +
    geom_jitter(data  = jitter_df, aes(y = isc_mean),
                width = 0.12, height = 0, size = 1.8, alpha = 0.5, colour = "grey55") +
    geom_errorbar(aes(ymin = mean_isc - se_isc, ymax = mean_isc + se_isc),
                  width = 0.12, linewidth = 0.8) +
    geom_point(size = 4) +
    scale_y_continuous(breaks = seq(0, 1, 0.1)) +
    labs(title = plot_title, x = NULL, y = "ISC (Pearson r)") +
    isc_theme
}
```

## 5.1 Version A — jitter by subject

```{r plot-subj, fig.width=5, fig.height=5}
dot_plot(subj_stats, subj_avg, "Inter-subject correlation\n(dots = subjects)")
ggsave(here("figures/isc_subjects.png"))
```

## 5.2 Version B — jitter by video

```{r plot-video, fig.width=10, fig.height=10, dpi=300}
  # Shared jitter position so each video's error line lines up with its dot
  video_jit <- position_jitter(width = 0.12, height = 0, seed = 42)

  # ── Category → color family, so clips from the same category (Pixar /
  # Slow / Sesame) render as shades of the same hue instead of unrelated
  # colors from the default discrete palette. Ramps run light -> dark
  # within a category so individual clips stay distinguishable. ────────────
  video_category_b <- function(v) {
    case_when(
      v == "pixar_birds"       ~ "Pixar",
      str_starts(v, "slow_")   ~ "Slow",
      str_starts(v, "sesame")  ~ "Sesame",
      str_starts(v, "frank")   ~ "Frank",
      TRUE                     ~ "Other"
    )
  }

  family_ramp <- list(
    Pixar  = colorRampPalette(c("#74c476", "#00441b")),  # greens
    Slow   = colorRampPalette(c("#9ecae1", "#08306b")),  # blues
    Sesame = colorRampPalette(c("#fa9fb5", "#7a0177")),  # pinks
    Frank  = colorRampPalette(c("#fdae6b", "#8c2d04")),  # oranges
    Other  = colorRampPalette(c("#cccccc", "#333333"))   # grey, fallback
  )

  video_color_df <- tibble(video_name = sort(unique(video_avg$video_name))) |>
    mutate(category = video_category_b(video_name)) |>
    arrange(category, video_name) |>
    group_by(category) |>
    mutate(shade_rank = row_number(), n_in_cat = n()) |>
    ungroup() |>
    rowwise() |>
    mutate(color = family_ramp[[category]](n_in_cat)[shade_rank]) |>
    ungroup()

  video_color_pal <- setNames(video_color_df$color, video_color_df$video_name)

  ggplot(video_stats, aes(x = pair_type, y = mean_isc)) +
    geom_linerange(data = video_avg,
                    aes(x = pair_type, y = isc_mean,
                        ymin = isc_mean - se_isc, ymax = isc_mean + se_isc,
                        color = video_name),
                    position = video_jit, alpha = 0.5, inherit.aes = FALSE) +
    geom_jitter(data  = video_avg, aes(y = isc_mean, color=video_name),
                position = video_jit, size = 4, alpha = 0.5) +
    geom_errorbar(aes(ymin = mean_isc - se_isc, ymax = mean_isc + se_isc),
                  width = 0.12, linewidth = 0.8) +
    geom_point(size = 4) +
    scale_color_manual(values = video_color_pal) +
    scale_y_continuous(breaks = seq(0, 1, 0.1)) +
    labs(title = "ISC by pair type\n(dots = videos)", x = NULL, y = "ISC (Pearson r)") +
    theme_few(base_size = 13) +
    theme(axis.text.x = element_text(size = 11),
          panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/isc_videos.png"))
```

## 5.2b Child–Child vs. Adult–Adult, by video

Three stacked panels, all sharing the same video order (grouped by video
category — Slow, Sesame, Frank, Pixar in that order — sorted ascending by
the Child–Child minus Adult–Adult difference within each group) and the
same per-video colors as 5.2 above. **Top**: Child–Child mean ISC.
**Middle**: Adult–Adult mean ISC (shares the top panel's y-axis range, for
direct comparison). **Bottom**: the difference between them. Error bars
are ± 1 SE (top/middle: `se_isc`; bottom: `sqrt(cc_se^2 + aa_se^2)`, valid
since the two groups are independent samples).

```{r plot-cc-aa-diff, fig.width=9, fig.height=12}
cc <- video_avg |>
  filter(pair_type == "Child–Child") |>
  select(video_name, cc_mean = isc_mean, cc_se = se_isc)

aa <- video_avg |>
  filter(pair_type == "Adult–Adult") |>
  select(video_name, aa_mean = isc_mean, aa_se = se_isc)

cat_order <- c("Slow", "Sesame", "Frank", "Pixar")

cc_aa_diff <- inner_join(cc, aa, by = "video_name") |>
  mutate(
    diff     = cc_mean - aa_mean,
    diff_se  = sqrt(cc_se^2 + aa_se^2),
    category = factor(video_category_b(video_name), levels = cat_order)
  ) |>
  arrange(category, diff) |>
  mutate(video_name = factor(video_name, levels = video_name))

# Shared y-axis range for the top (Child–Child) and middle (Adult–Adult)
# panels, so bar heights are visually comparable across panels.
isc_ylim <- range(
  c(cc_aa_diff$cc_mean - cc_aa_diff$cc_se, cc_aa_diff$cc_mean + cc_aa_diff$cc_se,
    cc_aa_diff$aa_mean - cc_aa_diff$aa_se, cc_aa_diff$aa_mean + cc_aa_diff$aa_se),
  na.rm = TRUE
)
isc_ylim <- c(0, isc_ylim[2] * 1.05)

no_x_axis <- theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

p_cc <- ggplot(cc_aa_diff, aes(x = video_name, y = cc_mean, fill = video_name)) +
  geom_col(width = 0.7, alpha = 0.9) +
  geom_errorbar(aes(ymin = cc_mean - cc_se, ymax = cc_mean + cc_se),
                width = 0.15, linewidth = 0.8) +
  scale_fill_manual(values = video_color_pal) +
  coord_cartesian(ylim = isc_ylim) +
  labs(title = "Child–Child ISC", x = NULL, y = "ISC (Pearson r)") +
  theme_few(base_size = 13) +
  theme(legend.position = "none", panel.grid.major.y = element_line(colour = "grey92")) +
  no_x_axis

p_aa <- ggplot(cc_aa_diff, aes(x = video_name, y = aa_mean, fill = video_name)) +
  geom_col(width = 0.7, alpha = 0.9) +
  geom_errorbar(aes(ymin = aa_mean - aa_se, ymax = aa_mean + aa_se),
                width = 0.15, linewidth = 0.8) +
  scale_fill_manual(values = video_color_pal) +
  coord_cartesian(ylim = isc_ylim) +
  labs(title = "Adult–Adult ISC", x = NULL, y = "ISC (Pearson r)") +
  theme_few(base_size = 13) +
  theme(legend.position = "none", panel.grid.major.y = element_line(colour = "grey92")) +
  no_x_axis

p_diff <- ggplot(cc_aa_diff, aes(x = video_name, y = diff, fill = video_name)) +
  geom_col(width = 0.7, alpha = 0.9) +
  geom_errorbar(aes(ymin = diff - diff_se, ymax = diff + diff_se),
                width = 0.15, linewidth = 0.8) +
  geom_hline(yintercept = 0, linewidth = 0.4, colour = "grey40") +
  scale_fill_manual(values = video_color_pal) +
  labs(title = "Child–Child minus Adult–Adult ISC",
       x = NULL, y = "ISC difference (Child–Child − Adult–Adult)") +
  theme_few(base_size = 13) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey92"))

p_cc / p_aa / p_diff
ggsave(here("figures/isc_cc_minus_aa.png"), width = 9, height = 12, dpi = 300)
```

## 5.3 Version C — bar plot by video category (Pixar vs. Slow vs. Sesame vs. Frank)

Same three pairings as above, but split by video category instead of
averaged over all clips: **Pixar** (`pixar_birds`) vs. **Slow** (the four
`slow_*` clips, averaged together) vs. **Sesame** (the four `sesame*`
clips, averaged together) vs. **Frank** (the three 'frank*' clips, averaged
together). Bars show the mean across subjects; dots are
individual participants' subject-level mean ISC for that category.

```{r build-plot-data-cat}
# ── Video → category lookup ────────────────────────────────────────────────
video_category <- function(v) {
  case_when(
    v == "pixar_birds"       ~ "Pixar",
    str_starts(v, "slow_")   ~ "Slow",
    str_starts(v, "sesame")  ~ "Sesame",
    str_starts(v, "frank")   ~ "Frank",
    TRUE                     ~ NA_character_
  )
}

isc_long_cat <- isc_long |>
  mutate(video_category = video_category(video_name)) |>
  filter(!is.na(video_category))

# ── Per-subject mean ISC, by pair type × video category ───────────────────
subj_avg_cat <- bind_rows(
  isc_long_cat |>
    filter(focal_group == "adults", partner_group == "adults") |>
    group_by(focal, video_category) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Adult–Adult"),

  isc_long_cat |>
    filter(focal_group == "kids", partner_group == "kids") |>
    group_by(focal, video_category) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Child–Child"),

  isc_long_cat |>
    filter(focal_group == "kids", partner_group == "adults") |>
    group_by(focal, video_category) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Child–Adult")
) |>
  mutate(pair_type = factor(pair_type, levels = plot_comparisons))

# ── Bar heights: mean ± SE across subjects ─────────────────────────────────
cat_stats <- subj_avg_cat |>
  group_by(pair_type, video_category) |>
  summarise(
    mean_isc = mean(isc_mean, na.rm = TRUE),
    se_isc   = sd(isc_mean, na.rm = TRUE) / sqrt(sum(!is.na(isc_mean))),
    .groups  = "drop"
  )
```

```{r plot-cat, fig.width=8, fig.height=6}
dodge_w <- 0.8
cat_jit <- position_jitterdodge(jitter.width = 0.15, dodge.width = dodge_w, seed = 42)

ggplot(cat_stats, aes(x = video_category, y = mean_isc, fill = pair_type)) +
  geom_col(position = position_dodge(width = dodge_w), width = 0.7, alpha = 0.8) +
  geom_jitter(data = subj_avg_cat, aes(y = isc_mean, group = pair_type),
              position = cat_jit, size = 1.6, alpha = 0.5, colour = "grey25") +
  geom_errorbar(aes(ymin = mean_isc - se_isc, ymax = mean_isc + se_isc),
                position = position_dodge(width = dodge_w), width = 0.15, linewidth = 0.8) +
  scale_fill_few() +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(title = "ISC by video category and pair type\n(dots = subjects)",
       x = NULL, y = "ISC (Pearson r)", fill = "Pair type") +
  theme_few(base_size = 13) +
  theme(axis.text.x = element_text(size = 11),
        panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/isc_category.png"))
```

---

# Part 6: Calibration Accuracy vs. ISC

Sanity-check plot: does calibration accuracy going into a block predict how
well a participant correlates with their same-age partners on that block's
videos? Calibration accuracy isn't stored anywhere alongside the ISC output
(fixation detection uses it once, to decide inclusion, then discards it), so
this recomputes it directly from each participant's own `trial_order.csv` /
`validation_summary.csv` -- same block-calibration logic as
`i2mc_fixations.py`'s `compute_block_calibration`, ported to R, including its
"an unknown calibration is not a known-good one" handling: a missing
`validation_summary.csv`, or a validation attempt whose accuracy failed to
compute (NaN), both come back `NA` here rather than silently joining as if
they'd passed.

```{r calibration-helpers}
# ── Per-participant calibration lookup (mirrors i2mc_fixations.py) ─────────────
find_cal_files <- function(pid, group) {
  pdir <- file.path(SERVER_PATH, "data/raw", group, pid)
  files <- list.files(pdir, full.names = TRUE)
  to <- files[str_detect(basename(files), "_trial_order\\.csv$")]
  vs <- files[str_detect(basename(files), "_validation_summary\\.csv$")]
  list(
    trial_order        = if (length(to) > 0) to[1] else NA_character_,
    validation_summary  = if (length(vs) > 0) vs[1] else NA_character_
  )
}

# Mean calibration accuracy (degrees) per block_index, from the last
# validation attempt before that block started -- same "last attempt by
# number, not last successfully-computed attempt" semantics as the Python
# version (see i2mc_fixations.py for why: a later attempt superseding an
# earlier one is intentional, even if that later attempt came back NaN).
compute_block_calibration_r <- function(trial_order_df, validation_summary_df) {
  mean_rows <- validation_summary_df |> filter(point == "mean")
  val_mean_deg <- setNames(
    (mean_rows$Mean_accuracy_degrees_left + mean_rows$Mean_accuracy_degrees_right) / 2,
    mean_rows$validation_step
  )

  first_trial_of_block <- trial_order_df |>
    group_by(block_index) |>
    summarise(first_trial = min(total_trial_index), .groups = "drop")

  block_cal <- setNames(
    rep(NA_real_, nrow(first_trial_of_block)),
    as.character(first_trial_of_block$block_index)
  )

  for (i in seq_len(nrow(first_trial_of_block))) {
    block_idx   <- first_trial_of_block$block_index[i]
    first_trial <- first_trial_of_block$first_trial[i]
    key <- as.character(block_idx)

    if (block_idx == 0) {
      pre_names <- sort(names(val_mean_deg)[str_starts(names(val_mean_deg), "pre_validation")])
      if (length(pre_names) > 0) block_cal[key] <- val_mean_deg[[tail(pre_names, 1)]]
      next
    }

    pat <- glue("^block_validation_trial{first_trial}_(\\d+)$")
    hits <- names(val_mean_deg)[str_detect(names(val_mean_deg), pat)]
    if (length(hits) > 0) {
      attempt_n <- as.integer(str_match(hits, pat)[, 2])
      block_cal[key] <- val_mean_deg[[hits[which.max(attempt_n)]]]
    }
  }
  block_cal
}

# calibration_unknown case: no validation_summary at all, or it's missing the
# expected columns -- NA, not "assume it passed" (matches the
# i2mc_fixations.py fix, not the pre-fix behavior).
get_block_calibration_deg <- function(pid, group, video_name) {
  cf <- find_cal_files(pid, group)
  if (is.na(cf$trial_order) || is.na(cf$validation_summary)) return(NA_real_)

  trial_order  <- suppressMessages(read_csv(cf$trial_order, show_col_types = FALSE))
  val_summary  <- suppressMessages(read_csv(cf$validation_summary, show_col_types = FALSE))
  if (!"point" %in% names(val_summary)) return(NA_real_)

  row <- trial_order |> filter(video_name == !!video_name)
  if (nrow(row) == 0) return(NA_real_)
  block_idx <- row$block_index[1]

  block_cal <- compute_block_calibration_r(trial_order, val_summary)
  key <- as.character(block_idx)
  if (!key %in% names(block_cal)) return(NA_real_)
  block_cal[[key]]
}
```

```{r calibration-compute}
# One representative video per (participant, video_category) -- any video
# within a category maps to the same block, so any of the participant's
# actual videos in isc_long_cat works to look up the block index.
cal_lookup <- isc_long_cat |>
  distinct(focal, focal_group, video_category, video_name) |>
  group_by(focal, focal_group, video_category) |>
  slice(1) |>
  ungroup() |>
  rowwise() |>
  mutate(calibration_deg = get_block_calibration_deg(focal, focal_group, video_name)) |>
  ungroup()

# Join against each participant's own within-group mean ISC for that
# category (Adult–Adult for adults, Child–Child for kids) -- same quantity
# as ISCw, just split by video_category instead of averaged across all clips.
same_group_isc <- subj_avg_cat |>
  filter(pair_type %in% c("Adult–Adult", "Child–Child")) |>
  mutate(focal_group = if_else(pair_type == "Adult–Adult", "adults", "kids"))

cal_isc <- cal_lookup |>
  inner_join(same_group_isc, by = c("focal", "focal_group", "video_category")) |>
  filter(!is.na(calibration_deg))

cat(glue("{nrow(cal_isc)} participant x category rows with a verified calibration reading ",
         "({n_distinct(cal_lookup$focal) - n_distinct(cal_isc$focal)} participants dropped for ",
         "unknown/unverifiable calibration)\n\n"))

cal_isc |>
  group_by(video_category, focal_group) |>
  summarise(
    r = round(cor(calibration_deg, isc_mean, use = "complete.obs"), 3),
    n = n(),
    .groups = "drop"
  )
```

```{r plot-calibration, fig.width=9, fig.height=5}
ggplot(cal_isc, aes(x = calibration_deg, y = isc_mean, colour = focal_group)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
  facet_wrap(~video_category) +
  scale_colour_few() +
  labs(title = "Calibration accuracy vs. mean within-group ISC",
       x = "Calibration accuracy (degrees, block for that video category)",
       y = "Mean ISC with same-age partners",
       colour = "Group") +
  theme_few(base_size = 13)
ggsave(here("figures/isc_calibration.png"))
```

---

# Part 7: ISC Timecourse (`r example_video`)

Sliding-window ISC over time for all three pairings, on `pixar_birds`. Unlike
every plot above, this can't be built from `isc_fixation_pairwise.csv` --
that file only has one aggregate ISC value per pair per video, no time
resolution. So this section recomputes windowed correlations directly from
each participant's fixation CSV: same overlap-weighted binning
(`fixations_to_binned_series`) and vectorized sliding-window Pearson
(`_pearson_matrix`) as `calc_isc_fixations.py` / `run_all_clips.py`, ported to R.
Only meaningful for `isc_source == "fixations"` (raw-gaze ISC has no
equivalent per-fixation file to bin from block calibration).

```{r timecourse-config}
# Must match the fixation code that produced isc_pairwise_csv above --
# not re-derived automatically, since the pairwise CSV doesn't carry it.
FIX_DIR   <- file.path(SERVER_PATH, "data/preprocessed/fixations")
FIX_CODE  <- "f329476c"
BIN_MS    <- 20
WINDOW_S  <- 2
STEP_S    <- 0.25
MIN_VALID <- 30   # bins required within a window for that pair's r to count
```

```{r timecourse-helpers}
# ── overlap-weighted winner-take-all binning, ported from
# calc_isc_fixations.py's fixations_to_binned_series: each bin gets the position
# of whichever fixation covers the largest share of that bin's duration, so
# a bin straddling a fixation boundary is resolved by overlap instead of by
# which fixation happened to be read last. ──────────────────────────────────
fixations_to_binned_series_r <- function(fix_df, bin_ms) {
  if (nrow(fix_df) == 0) return(list(x = numeric(0), y = numeric(0)))
  n_bins <- as.integer(max(fix_df$endT) %/% bin_ms) + 1L
  x <- rep(NA_real_, n_bins)
  y <- rep(NA_real_, n_bins)
  best_overlap <- rep(0, n_bins)

  for (i in seq_len(nrow(fix_df))) {
    startT <- fix_df$startT[i]; endT <- fix_df$endT[i]
    b0 <- as.integer(startT %/% bin_ms); b1 <- as.integer(endT %/% bin_ms)
    for (b in b0:b1) {
      bin_start <- b * bin_ms; bin_end <- (b + 1) * bin_ms
      overlap <- min(endT, bin_end) - max(startT, bin_start)
      if (overlap > best_overlap[b + 1]) {
        best_overlap[b + 1] <- overlap
        x[b + 1] <- fix_df$xpos[i]
        y[b + 1] <- fix_df$ypos[i]
      }
    }
  }
  list(x = x, y = y)
}

load_binned <- function(pid, group, video, code, bin_ms = BIN_MS) {
  path <- file.path(FIX_DIR, group, glue("{pid}_{video}_{code}.csv"))
  if (!file.exists(path)) return(NULL)
  fix_df <- suppressMessages(read_csv(path, show_col_types = FALSE))
  fixations_to_binned_series_r(fix_df, bin_ms)
}

# ── vectorized windowed Pearson r for N pairs simultaneously, ported from
# run_all_clips.py's _pearson_matrix -- computes all pairs in one window in
# a single matrix operation instead of looping pair-by-pair. ───────────────
pearson_matrix <- function(mat_a, mat_b, min_valid = MIN_VALID) {
  valid <- !is.na(mat_a) & !is.na(mat_b)
  n <- rowSums(valid)
  a <- ifelse(valid, mat_a, 0)
  b <- ifelse(valid, mat_b, 0)
  sum_a <- rowSums(a); sum_b <- rowSums(b)
  sum_aa <- rowSums(a * a); sum_bb <- rowSums(b * b); sum_ab <- rowSums(a * b)
  num <- n * sum_ab - sum_a * sum_b
  denom <- sqrt(pmax((n * sum_aa - sum_a^2) * (n * sum_bb - sum_b^2), 0))
  ok <- (n >= min_valid) & (denom > 1e-10)
  ifelse(ok, num / denom, NA_real_)
}
```

```{r timecourse-peak-fn}
# ── Contiguous high-ISC periods, by comparison ─────────────────────────────
# "Highest" = top 30% of that comparison's own windows (70th percentile
# threshold on mean_r, computed separately per comparison -- adults' and
# kids' thresholds are not the same r value); a period is a run of
# consecutive 1s-step windows that all clear that threshold, kept only if
# it spans >= 2s between the first and last qualifying window. 0.70 was
# chosen empirically for Child–Child on pixar_birds -- it's the quantile in
# a sweep from the 50th to 85th percentile that landed closest to the middle
# of the ~7-10 periods wanted, using the boundary convention below (a
# tighter convention needs a looser quantile to find the same number of
# periods, since fewer runs are long enough to survive); the same quantile
# is applied to every comparison and every video below as "the same
# approach", not independently re-tuned per video to also land in 7-10.
#
# Boundary convention: end_s is the LAST qualifying window's own start_s,
# not start_s + window_s. Adding window_s would extend the shaded region
# past the last confirmed-high point into the next (non-qualifying, often
# already-dropped) window's territory -- correct in the narrow sense that
# a window covering [t, t+window_s] being high does say something about
# ISC near t+window_s, but visually indefensible: it draws shading over a
# stretch of the line that's often already visibly dropped, since the line
# itself only has vertices at window_start_s and doesn't show that
# reach. Keeping both edges anchored to actual qualifying vertices (same
# convention start_s already used) makes the shading match what the line
# actually shows, at the cost of a stricter effective duration filter
# (duration is now (n-1)*STEP_S, not (n-1)*STEP_S + window_s -- a run needs
# one more qualifying window to clear the same min_dur_s).
PEAK_QUANTILE  <- 0.70
PEAK_MIN_DUR_S <- 1.5

find_high_isc_periods <- function(windowed_df, comparison_label, quantile = PEAK_QUANTILE,
                                   min_dur_s = PEAK_MIN_DUR_S) {
  series <- windowed_df |> filter(comparison == comparison_label) |> arrange(window_start_s)
  threshold <- quantile(series$mean_r, quantile, na.rm = TRUE)

  above <- series$mean_r >= threshold
  run <- rle(above)
  run_ends <- cumsum(run$lengths)
  run_starts <- run_ends - run$lengths + 1

  periods <- tibble(run_start_idx = run_starts, run_end_idx = run_ends, is_high = run$values) |>
    filter(is_high) |>
    mutate(
      start_s  = series$window_start_s[run_start_idx],
      end_s    = series$window_start_s[run_end_idx],
      duration = end_s - start_s
    ) |>
    filter(duration >= min_dur_s) |>
    select(start_s, end_s, duration)

  attr(periods, "threshold") <- threshold
  periods
}
```

```{r timecourse-build-fn}
# ── One function, any video: load fixation data for every adult/kid who has
# it for `video`, compute windowed ISC for all three pairings, detect
# high-ISC periods, build the plot. Used once below for `example_video` and
# again, in a loop, for the four slow_* videos -- same pipeline either way,
# just parameterized by video name instead of hardcoded to one.
## bin_ms/window_s/step_s/min_valid/valid_adult_pids default to the globals configured
## above, so every EXISTING call site (build_timecourse(.x), build_timecourse(example_video,
## title=...)) is completely unaffected -- these params exist only for the
## concordance-comparison export in "Part 7c" below, which calls this with different values.
build_timecourse <- function(video_name, title = video_name, base_size = 13,
                              bin_ms = BIN_MS, window_s = WINDOW_S, step_s = STEP_S,
                              min_valid = MIN_VALID, valid_adult_pids = NULL) {
  pix <- isc_raw |> filter(video_name == !!video_name)
  adults <- sort(unique(c(pix$pid_a[pix$group_a == "adults"], pix$pid_b[pix$group_b == "adults"])))
  if (!is.null(valid_adult_pids)) adults <- intersect(adults, valid_adult_pids)
  kids   <- sort(unique(c(pix$pid_a[pix$group_a == "kids"],   pix$pid_b[pix$group_b == "kids"])))

  pids      <- c(adults, kids)
  groups    <- c(rep("adults", length(adults)), rep("kids", length(kids)))
  groups_of <- set_names(groups, pids)
  idx_of    <- set_names(seq_along(pids), pids)

  binned <- map2(pids, groups, load_binned, video = video_name, code = FIX_CODE, bin_ms = bin_ms)
  names(binned) <- pids

  # Median, not max, sizes the shared bin matrix -- a single participant
  # whose trial data runs past the video's actual end (seen once so far:
  # one kid's slow_hands file extends ~14s beyond the real ~60.7s video,
  # while every other participant/video combination clusters within ~30ms
  # of the median) would otherwise single-handedly stretch every panel's
  # x-axis to match their bad data. Participants longer than the median are
  # truncated to fit; this only discards time past where virtually everyone
  # else's trial already ended, not real mid-video data.
  bin_lengths <- map_int(binned, ~ length(.x$x))
  n_bins <- as.integer(round(median(bin_lengths)))
  X <- matrix(NA_real_, nrow = length(pids), ncol = n_bins)
  Y <- matrix(NA_real_, nrow = length(pids), ncol = n_bins)
  for (i in seq_along(pids)) {
    b <- binned[[i]]
    len_i <- min(length(b$x), n_bins)
    X[i, seq_len(len_i)] <- b$x[seq_len(len_i)]
    Y[i, seq_len(len_i)] <- b$y[seq_len(len_i)]
  }

  pair_group <- function(pairs, label_a, label_b) {
    keep(pairs, function(p) {
      ga <- groups_of[p[1]]; gb <- groups_of[p[2]]
      (ga == label_a && gb == label_b) || (ga == label_b && gb == label_a)
    })
  }
  all_pairs <- combn(pids, 2, simplify = FALSE)
  aa_pairs <- pair_group(all_pairs, "adults", "adults")
  kk_pairs <- pair_group(all_pairs, "kids", "kids")
  ak_pairs <- pair_group(all_pairs, "adults", "kids")

  window_bins <- as.integer(round(window_s * 1000 / bin_ms))
  step_bins   <- as.integer(round(step_s * 1000 / bin_ms))
  n_windows   <- max(0, (n_bins - window_bins) %/% step_bins + 1)

  windowed_r_for_pairs <- function(pairs) {
    ia <- map_int(pairs, ~ idx_of[[.x[1]]])
    ib <- map_int(pairs, ~ idx_of[[.x[2]]])
    out <- vector("list", n_windows)
    for (w in seq_len(n_windows)) {
      start <- (w - 1) * step_bins + 1
      end <- start + window_bins - 1
      r_x <- pearson_matrix(X[ia, start:end, drop = FALSE], X[ib, start:end, drop = FALSE], min_valid = min_valid)
      r_y <- pearson_matrix(Y[ia, start:end, drop = FALSE], Y[ib, start:end, drop = FALSE], min_valid = min_valid)
      r_avg <- rowMeans(cbind(r_x, r_y), na.rm = TRUE)
      r_avg[is.nan(r_avg)] <- NA_real_
      out[[w]] <- tibble(window_start_s = (start - 1) * bin_ms / 1000, r = r_avg)
    }
    bind_rows(out)
  }

  windowed <- bind_rows(
    windowed_r_for_pairs(aa_pairs) |> mutate(comparison = "Adult–Adult"),
    windowed_r_for_pairs(kk_pairs) |> mutate(comparison = "Child–Child"),
    windowed_r_for_pairs(ak_pairs) |> mutate(comparison = "Child–Adult")
  ) |>
    group_by(comparison, window_start_s) |>
    summarise(mean_r = mean(r, na.rm = TRUE), n_pairs = sum(!is.na(r)), .groups = "drop")

  kk_peaks <- find_high_isc_periods(windowed, "Child–Child")
  aa_peaks <- find_high_isc_periods(windowed, "Adult–Adult")

  p <- ggplot(windowed, aes(x = window_start_s, y = mean_r)) +
    geom_rect(data = kk_peaks, aes(xmin = start_s, xmax = end_s, ymin = -Inf, ymax = Inf),
              inherit.aes = FALSE, fill = "grey70", colour = NA, alpha = 0.4) +
    geom_rect(data = aa_peaks, aes(xmin = start_s, xmax = end_s, ymin = -Inf, ymax = Inf),
              inherit.aes = FALSE, fill = NA, colour = "black", linewidth = 0.5) +
    geom_line(aes(colour = comparison), linewidth = 0.8) +
    scale_colour_few() +
    scale_x_continuous(breaks = seq(0, max(windowed$window_start_s), by = 5),
                        minor_breaks = seq(0, max(windowed$window_start_s), by = 1),
                        guide = guide_axis(minor.ticks = TRUE)) +
    labs(title = title, x = "Time (s)", y = "Mean ISC (r)", colour = "Comparison") +
    theme_few(base_size = base_size) +
    theme(panel.grid.minor.x = element_blank(),
          axis.minor.ticks.x.bottom = element_line(colour = "grey30", linewidth = 0.3),
          axis.minor.ticks.length = unit(3, "pt"))

  list(plot = p, windowed = windowed, kk_peaks = kk_peaks, aa_peaks = aa_peaks,
       n_adults = length(adults), n_kids = length(kids), n_bins = n_bins)
}
```

```{r plot-timecourse, fig.width=18, fig.height=3}
tc_pixar <- build_timecourse(
  example_video,
  title = glue("Fixation-position ISC over time — {example_video}")
)
cat(glue("{tc_pixar$n_adults} adults, {tc_pixar$n_kids} kids, {tc_pixar$n_bins} bins ",
         "({round(tc_pixar$n_bins * BIN_MS / 1000, 1)}s) at {BIN_MS}ms resolution\n"))
cat(glue("{nrow(tc_pixar$kk_peaks)} high-Child–Child-ISC periods (threshold r >= ",
         "{round(attr(tc_pixar$kk_peaks, 'threshold'), 3)}); ",
         "{nrow(tc_pixar$aa_peaks)} high-Adult–Adult-ISC periods (threshold r >= ",
         "{round(attr(tc_pixar$aa_peaks, 'threshold'), 3)})\n"))

tc_pixar$plot +
  labs(subtitle = glue("{WINDOW_S}s sliding window, {STEP_S}s step, {BIN_MS}ms bins — shaded: top ",
                        "{round((1 - PEAK_QUANTILE) * 100)}% periods, >= {PEAK_MIN_DUR_S}s ",
                        "(grey fill = Child–Child, black border = Adult–Adult)"))
ggsave(here("figures/isc_timecourse.png"), width = 18, height = 3)
```

---

# Part 7b: ISC Timecourse — `slow_*` videos

Same pipeline as Part 7, run once per `slow_*` video and stacked into one
4-panel figure (one row per video) instead of four separate plots -- makes
it easy to see that these four clips behave quite differently from each
other (recall `slow_people` had much higher whole-trial ISC than
`slow_inanimate`/`slow_animals` in the Version C bar plot above).

```{r timecourse-slow-build}
slow_videos <- c("slow_animals", "slow_hands", "slow_inanimate", "slow_people")

slow_results <- map(slow_videos, ~ build_timecourse(.x, title = .x, base_size = 11))
names(slow_results) <- slow_videos

walk(slow_videos, function(v) {
  r <- slow_results[[v]]
  cat(glue("{v}: {r$n_adults} adults, {r$n_kids} kids -- ",
           "{nrow(r$kk_peaks)} Child–Child / {nrow(r$aa_peaks)} Adult–Adult high-ISC periods\n"))
})
```

```{r plot-timecourse-slow, fig.width=9, fig.height=10}
slow_plots <- map(slow_results, "plot")
wrap_plots(slow_plots, ncol = 1) +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Fixation-position ISC over time — slow_* videos",
    subtitle = glue("{WINDOW_S}s sliding window, {STEP_S}s step, {BIN_MS}ms bins — shaded: top ",
                     "{round((1 - PEAK_QUANTILE) * 100)}% periods, >= {PEAK_MIN_DUR_S}s ",
                     "(grey fill = Child–Child, black border = Adult–Adult)")
  )
ggsave(here("figures/isc_timecourse_slow.png"), width = 9, height = 10)
```

---

# Part 8: ISC Timecourse Shaded by Shot Scale (All Videos)

Same windowed fixation-position ISC as Part 7 (2s window, 0.25s step, all
three pairings) -- reuses `build_timecourse()`'s `windowed` output directly,
no recomputation -- but instead of shading/bordering just the high-ISC
periods, the *entire* plot background is shaded continuously by shot scale
(ECS/CS/MFS/LS -- see `analysis/stage1_shot_scale.py` and
`annotations/shot_scale/` for how these were produced: VideoMAE shot-scale
classifier, 0.667s non-overlapping tiles, three-crop majority/median
resolution). This is a pure visual overlay -- the shot-scale tiles (0.667s
wide) and the ISC windows (2s wide) sit on two independent time grids here;
nothing is resampled or jointly resolved onto a shared grid for this plot,
that's a separate question for actual analysis, not visualization.

Covers all 12 stimulus videos (Part 7/7b covered only `pixar_birds` and the
`slow_*` clips).

```{r shotscale-setup}
SHOTSCALE_DIR    <- here("annotations/shot_scale")
SHOTSCALE_LEVELS <- c("ECS", "CS", "MFS", "LS")

# Sequential ramp (ordinal scale, close -> far), kept pale throughout so the
# ISC lines and gridlines stay legible drawn on top.
SHOTSCALE_COLORS <- setNames(
  colorRampPalette(c("#f7fbff", "#6baed6"))(length(SHOTSCALE_LEVELS)),
  SHOTSCALE_LEVELS
)

read_shotscale <- function(video_name) {
  path <- Sys.glob(here(glue("annotations/shot_scale/{video_name}_shotscale_0-*s.csv")))[1]
  read_csv(path, show_col_types = FALSE) |>
    mutate(resolved_class = factor(resolved_class, levels = SHOTSCALE_LEVELS))
}
```

```{r shotscale-build-fn}
# Takes a precomputed `windowed` df (from build_timecourse(video_name)$windowed)
# rather than recomputing it -- the caller loops over all 12 videos, and Part 9
# needs the same per-video windowed timecourses again, so both reuse one shared
# cache instead of each independently paying for 12 calls to build_timecourse().
build_timecourse_shotscale <- function(video_name, windowed, title = video_name, base_size = 13) {
  shot <- read_shotscale(video_name)

  ggplot() +
    geom_rect(data = shot, aes(xmin = start_s, xmax = end_s, ymin = -Inf, ymax = Inf, fill = resolved_class),
              inherit.aes = FALSE, alpha = 0.9) +
    scale_fill_manual(values = SHOTSCALE_COLORS, name = "Shot scale", drop = FALSE) +
    geom_line(data = windowed, aes(x = window_start_s, y = mean_r, colour = comparison), linewidth = 0.8) +
    scale_colour_few() +
    scale_x_continuous(breaks = seq(0, max(windowed$window_start_s), by = 5),
                        minor_breaks = seq(0, max(windowed$window_start_s), by = 1),
                        guide = guide_axis(minor.ticks = TRUE)) +
    labs(title = title, x = "Time (s)", y = "Mean ISC (r)", colour = "Comparison") +
    theme_few(base_size = base_size) +
    theme(panel.grid.minor.x = element_blank(),
          axis.minor.ticks.x.bottom = element_line(colour = "grey30", linewidth = 0.3),
          axis.minor.ticks.length = unit(3, "pt"))
}
```

```{r shotscale-plot-all, fig.width=18, fig.height=3}
all_video_names <- sort(unique(isc$video_name))

# Computed once here, reused by Part 9 below -- avoids recomputing
# build_timecourse() (the expensive step) a second time for all 12 videos.
windowed_cache <- set_names(map(all_video_names, ~ build_timecourse(.x)$windowed), all_video_names)

for (v in all_video_names) {
  cat(glue("Building shot-scale-shaded timecourse for {v}...\n"))
  p <- build_timecourse_shotscale(v, windowed_cache[[v]], title = glue("Fixation-position ISC over time, shaded by shot scale — {v}"))
  print(p)
  ggsave(here(glue("figures/isc_timecourse_shotscale_{v}.png")), plot = p, width = 18, height = 3)
}
```

---

# Part 8b: Export Windowed ISC for Concordance Comparison

Writes a CSV of the windowed Adult–Adult (and Child–Child / Child–Adult) ISC for every
clip, so it can be read by the fixation-density-concordance R scripts/notebook
(`calc_concordance_group.R`, `calc_concordance_pairwise.R`,
`plot_concordance_isc_timecourses.qmd`) without re-deriving ISC from scratch in Python. This
supersedes `calc_isc_timecourse.py`/`isc_fixation_timecourse_1s.csv`, the earlier
Python port that existed only because this notebook didn't export a CSV yet.

Computed at DIFFERENT parameters than Part 7/7b/8/9's own `BIN_MS=20, WINDOW_S=2` (those
stay exactly as they were, untouched, via `build_timecourse`'s new params defaulting back
to the existing globals) -- concordance uses a 1.0s window, and `BIN_MS=25` instead of 20
here specifically because 20 doesn't divide 250ms evenly (rounds to a 240ms step, silently
misaligning this export's grid against concordance's exact 250ms grid -- see
METHODS_DECISIONS.md, 2026-10-01/02, for the full story of that bug and this fix).

**Participant population, matched to concordance (METHODS_DECISIONS.md, 2026-10-02):**
`isc_raw`'s own adult/pid list for a video (whoever has a row in the pre-computed
`isc_fixation_pairwise.csv`) does NOT equal the `valid_data==TRUE` population
`calc_concordance_group.R` uses (confirmed by testing: 44 vs. 41 adults for
`pixar_birds`, a 3/7-person mismatch in each direction) -- `isc_fixation_pairwise.csv`'s
own inclusion rule is just "has a fixation file for this code", with no `valid_data` QC
step. Filtering to `valid_adult_pids` here is REQUIRED for this export to be a fair
comparison against concordance, not optional.

```{r export-config}
EXPORT_BIN_MS    <- 25     # divides 250ms evenly (10 bins); BIN_MS=20 does not (rounds to 240ms)
EXPORT_WINDOW_S  <- 1.0    # matches calc_concordance_group.R's WINDOW_S
EXPORT_STEP_S    <- 0.25   # matches calc_concordance_group.R's STEP_S
EXPORT_MIN_VALID <- 30     # same bin-validity gate as Part 7's MIN_VALID; concordance has
                           # no equivalent minimum-data gate by design (METHODS_DECISIONS.md)

participant_summary <- read_csv(here("data", "local_cache", "participant_summary.csv"), show_col_types = FALSE)
valid_adult_pids <- participant_summary |>
  filter(participant_type == "adults", valid_data == TRUE) |>
  pull(participant_id)
cat(glue("{length(valid_adult_pids)} valid adults (matches concordance's population)\n"))
```

```{r export-compute}
export_cache <- set_names(
  map(all_video_names, ~ build_timecourse(
    .x,
    bin_ms = EXPORT_BIN_MS, window_s = EXPORT_WINDOW_S, step_s = EXPORT_STEP_S,
    min_valid = EXPORT_MIN_VALID, valid_adult_pids = valid_adult_pids
  )$windowed),
  all_video_names
)

isc_timecourse_export <- map_dfr(all_video_names, function(v) {
  export_cache[[v]] |>
    arrange(comparison, window_start_s) |>
    group_by(comparison) |>
    mutate(window_idx = row_number() - 1L,
           window_end_s = round(window_start_s + EXPORT_WINDOW_S, 3)) |>
    ungroup() |>
    mutate(video_name = v, window_start_s = round(window_start_s, 3)) |>
    select(video_name, comparison, window_idx, window_start_s, window_end_s, mean_r, n_pairs)
})

write_csv(isc_timecourse_export, here("data", "results", "concordance", "isc_fixation_timecourse.csv"))
cat(glue("Wrote {nrow(isc_timecourse_export)} rows ({length(all_video_names)} videos) to ",
         "data/results/concordance/isc_fixation_timecourse.csv\n"))
isc_timecourse_export |> filter(comparison == "Adult–Adult") |> count(video_name, name = "n_windows")
```

---

# Part 9: Mean ISC by Shot Scale

Reconciles the two independent time grids from Part 8 (2s/0.25s ISC windows
vs. 0.667s shot-scale tiles) so each ISC window can be pooled into one of
the four shot-scale bins. Broken out by comparison (Adult–Adult / Child–Adult
/ Child–Child) -- three bars per shot-scale class -- and, from here on, kept
**per video rather than pooled/scaled across videos**: one facet per video,
raw `mean_r`, no min-max rescaling. Pooling videos together (with or without
per-video scaling) was useful for a first look, but a shot-scale effect that
actually depends on video-specific content is exactly what pooling would
wash out, so this is the more conservative default going forward.

**Reconciliation rule**: each ISC window is assigned the shot-scale class of
whichever tile has the largest time-overlap with that window -- the same
"largest overlap wins" convention `fixations_to_binned_series` already uses
elsewhere in this file for an analogous binning problem, not a new ad hoc
rule.

```{r shotscale-reconcile}
# Adult–Adult / Child–Adult / Child–Child, in that order so Child–Adult --
# the comparison of most interest -- lands in the middle of each 3-bar
# dodge cluster.
COMPARISON_LEVELS <- c("Adult–Adult", "Child–Adult", "Child–Child")

# Vectorized per video: an n_windows x n_tiles overlap matrix, take
# whichever tile has the largest overlap for each window.
assign_shotscale_video <- function(window_starts, shot_df, window_s = WINDOW_S) {
  window_ends <- window_starts + window_s
  overlap <- outer(window_ends, shot_df$end_s, pmin) - outer(window_starts, shot_df$start_s, pmax)
  overlap[overlap < 0] <- 0
  best_tile <- apply(overlap, 1, which.max)
  best_overlap <- apply(overlap, 1, max)
  cls <- as.character(shot_df$resolved_class[best_tile])
  cls[best_overlap <= 0] <- NA_character_  # window has no overlapping tile at all
  cls
}

# Reuses the windowed timecourses Part 8 already computed (windowed_cache)
# instead of calling build_timecourse() again for all 12 videos.
windowed_shotscale <- map_dfr(all_video_names, function(v) {
  windowed <- windowed_cache[[v]]
  shot <- read_shotscale(v)
  windowed$shot_class <- assign_shotscale_video(windowed$window_start_s, shot)
  windowed$video_name <- v
  windowed
})

cat(glue("{sum(is.na(windowed_shotscale$shot_class))} / {nrow(windowed_shotscale)} ",
         "windows had no overlapping shot-scale tile (dropped from analysis below)\n"))
```

```{r shotscale-aggregate}
# Mean raw ISC per (video, shot_class, comparison) -- no cross-video pooling,
# no scaling.
video_class <- windowed_shotscale |>
  filter(!is.na(shot_class)) |>
  mutate(shot_class = factor(shot_class, levels = SHOTSCALE_LEVELS),
         comparison = factor(comparison, levels = COMPARISON_LEVELS)) |>
  group_by(video_name, shot_class, comparison) |>
  summarise(mean_raw = mean(mean_r, na.rm = TRUE), .groups = "drop")

print(video_class, n = Inf)
```

```{r plot-shotscale-bar, fig.width=18, fig.height=12}
ggplot(video_class, aes(x = shot_class, y = mean_raw, fill = comparison)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.65, alpha = 0.85) +
  scale_fill_few(name = "Comparison") +
  facet_wrap(~video_name, ncol = 4) +
  labs(title = "Mean ISC by shot scale, by video",
       subtitle = "Raw, unscaled -- one panel per video, no cross-video pooling",
       x = "Shot scale", y = "Mean ISC (raw r)") +
  theme_few(base_size = 12) +
  theme(panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/isc_by_shotscale.png"), width = 18, height = 12)
```

---

# Part 10: Fixation-Location Entropy by Shot Scale (data build)

The windowed ISC in Parts 7-9 turned out to need a long time series within
each window to detect a real effect (confirmed by comparing whole-trial ISC
against 2s-windowed ISC on the `slow_*` videos: pooled Adult-Adult 0.369 vs.
Child-Child 0.296 at the whole-trial level, vs. 0.150 vs. 0.149 -- basically
gone -- once windowed into 2s chunks). Shannon entropy of *where* fixations
land doesn't have that problem: it's a snapshot statistic over pooled
fixation locations, not a time-series correlation, so a short 0.667s
shot-scale tile is enough data to compute it from, as long as enough
fixations get pooled into it.

**No duration weighting.** This is a purely spatial analysis (dispersion of
*where* people looked, not *how long*); weighting by fixation duration would
reintroduce a temporal question this section is deliberately sidestepping.
Each fixation counts as one point.

**Tile assignment**: for each participant, each shot-scale tile gets that
participant's single fixation with the largest time-overlap with the tile --
the same "largest overlap wins" rule used everywhere else in this notebook
for this exact kind of boundary problem (`assign_shotscale_video`,
`fixations_to_binned_series_r`). A participant contributes no point to a
tile if none of their fixations overlap it at all.

**Grid**: 4x2 cells over the 1920x1080 display (480x540 px each), giving a
max possible entropy of log2(8) = 3 bits. Conceptually the same computation
as MATLAB's `entropy()` -- Shannon entropy of a normalized histogram -- just
applied directly to a 2D histogram of fixation coordinates instead of an
image's pixel-intensity histogram, so there's no rendering/smoothing/
quantization step in between the raw data and the number. Off-screen
fixations (coordinates outside the display) are excluded rather than
clamped to an edge cell, so eye-tracker overshoot doesn't inflate edge-cell
counts.

**Kids vs. adults, not three pairings.** Entropy is a within-group spatial
statistic, not a pairwise one, so there's no Child-Adult analog here --
just an adults series and a kids series per shot-scale tile.

**Rarefaction**: kids are the smaller group in every video (12-14 vs. up to
44 adults), so kid entropy is computed directly from all of that tile's
available kid points, unweighted. Adults are repeatedly subsampled down to
that video's kid count (500 draws without replacement, or fewer if fewer
adult points are actually available in that tile), with entropy recomputed
each draw; both the mean and SD across draws are kept, so the rarefaction
noise itself is visible rather than hidden behind a single arbitrary draw.

**Minimum count floor**: below 5 pooled points (post-matching), entropy is
too close to the degenerate "one point per bin" regime to trust, and that
cell is left `NA` rather than plotted as if it were a real estimate.

This section only builds and saves the long-format table -- one row per
(video, group, tile) -- deliberately without pooling across videos or
picking a plot yet, so those decisions can be made later by reading
`data/preprocessed/entropy_by_time.csv` directly instead of rerunning this
whole notebook.

```{r entropy-config}
ENTROPY_GRID_NX <- 4
ENTROPY_GRID_NY <- 2
SCREEN_W <- 1920
SCREEN_H <- 1080
ENTROPY_MIN_N   <- 5    # below this many pooled points, entropy for that (video, group, tile) is unreliable -> NA
ENTROPY_N_DRAWS <- 500  # rarefaction draws for adults, subsampled down to that video's kid N
set.seed(42)             # reproducible rarefaction, same convention as the I2MC k-means seed elsewhere in this project
```

```{r entropy-helpers}
# One (x, y) point per participant per tile: whichever of that participant's
# raw fixations has the largest time-overlap with the tile. Same convention
# as assign_shotscale_video() and fixations_to_binned_series_r() above --
# "largest overlap wins" is the one rule this notebook uses everywhere a
# fixation/window has to be assigned to a single bin.
assign_fixations_to_tiles <- function(fix_df, tile_start_s, tile_end_s) {
  if (nrow(fix_df) == 0) {
    return(tibble(tile_start_s = tile_start_s, tile_end_s = tile_end_s,
                  x = NA_real_, y = NA_real_))
  }
  fs <- fix_df$startT / 1000
  fe <- fix_df$endT / 1000
  overlap <- outer(tile_end_s, fe, pmin) - outer(tile_start_s, fs, pmax)
  overlap[overlap < 0] <- 0
  best_fix <- apply(overlap, 1, which.max)
  best_overlap <- apply(overlap, 1, max)
  x <- fix_df$xpos[best_fix]
  y <- fix_df$ypos[best_fix]
  x[best_overlap <= 0] <- NA_real_
  y[best_overlap <= 0] <- NA_real_
  tibble(tile_start_s = tile_start_s, tile_end_s = tile_end_s, x = x, y = y)
}

load_tile_points <- function(pid, group, video, tile_start_s, tile_end_s, code = FIX_CODE) {
  path <- file.path(FIX_DIR, group, glue("{pid}_{video}_{code}.csv"))
  if (!file.exists(path)) {
    return(tibble(tile_start_s = tile_start_s, tile_end_s = tile_end_s, x = NA_real_, y = NA_real_))
  }
  fix_df <- suppressMessages(read_csv(path, show_col_types = FALSE))
  assign_fixations_to_tiles(fix_df, tile_start_s, tile_end_s)
}

# Shannon entropy (bits) of a set of (x, y) points over a fixed grid --
# conceptually the same computation as MATLAB's entropy(), just on a 2D
# histogram of raw fixation coordinates instead of an image's pixel
# intensities. Off-screen points are dropped, not clamped to an edge cell.
shannon_entropy_bits <- function(x, y, nx = ENTROPY_GRID_NX, ny = ENTROPY_GRID_NY,
                                  screen_w = SCREEN_W, screen_h = SCREEN_H) {
  ok <- !is.na(x) & !is.na(y) & x >= 0 & x <= screen_w & y >= 0 & y <= screen_h
  x <- x[ok]; y <- y[ok]
  if (length(x) == 0) return(NA_real_)
  cx <- pmin(nx, floor(x / (screen_w / nx)) + 1)
  cy <- pmin(ny, floor(y / (screen_h / ny)) + 1)
  counts <- tabulate((cy - 1) * nx + cx, nbins = nx * ny)
  p <- counts[counts > 0] / sum(counts)
  -sum(p * log2(p))
}
```

```{r entropy-build}
tile_points <- map_dfr(all_video_names, function(v) {
  shot <- read_shotscale(v)
  pix <- isc_raw |> filter(video_name == !!v)
  adults <- sort(unique(c(pix$pid_a[pix$group_a == "adults"], pix$pid_b[pix$group_b == "adults"])))
  kids   <- sort(unique(c(pix$pid_a[pix$group_a == "kids"],   pix$pid_b[pix$group_b == "kids"])))

  participant_points <- function(pid, group) {
    load_tile_points(pid, group, v, shot$start_s, shot$end_s) |>
      mutate(pid = pid, group = group)
  }

  bind_rows(
    map_dfr(adults, participant_points, group = "adults"),
    map_dfr(kids,   participant_points, group = "kids")
  ) |>
    mutate(video_name = v, kid_n_video = length(kids)) |>
    left_join(shot |> select(start_s, end_s, resolved_class),
              by = c("tile_start_s" = "start_s", "tile_end_s" = "end_s")) |>
    rename(shot_class = resolved_class)
})

cat(glue("{sum(!is.na(tile_points$x))} / {nrow(tile_points)} participant x tile points had a valid overlapping fixation\n"))
```

```{r entropy-compute}
# Kids: deterministic entropy from ALL of that tile's available kid points --
# no subsampling, since kids are already the smaller group.
entropy_kids <- tile_points |>
  filter(group == "kids") |>
  group_by(video_name, tile_start_s, tile_end_s, shot_class) |>
  summarise(n_available = sum(!is.na(x)),
            n_used = n_available,
            entropy_bits = if_else(n_used >= ENTROPY_MIN_N,
                                    shannon_entropy_bits(x, y), NA_real_),
            entropy_sd_bits = NA_real_, n_draws = NA_integer_,
            .groups = "drop") |>
  mutate(group = "kids")

# Adults: rarefied -- repeatedly subsample down to that video's kid N (or
# however many adult points are actually available in this tile, if that's
# smaller), recompute entropy each draw, report the mean and SD across
# draws so the rarefaction noise itself is visible rather than hidden in a
# single arbitrary draw.
entropy_adults <- tile_points |>
  filter(group == "adults") |>
  group_by(video_name, tile_start_s, tile_end_s, shot_class, kid_n_video) |>
  group_modify(function(df, keys) {
    ok <- !is.na(df$x) & !is.na(df$y)
    xa <- df$x[ok]; ya <- df$y[ok]
    n_available <- length(xa)
    n_used <- min(keys$kid_n_video, n_available)
    if (n_used < ENTROPY_MIN_N) {
      return(tibble(n_available = n_available, n_used = n_used,
                     entropy_bits = NA_real_, entropy_sd_bits = NA_real_,
                     n_draws = NA_integer_))
    }
    draws <- replicate(ENTROPY_N_DRAWS, {
      idx <- sample.int(n_available, n_used, replace = FALSE)
      shannon_entropy_bits(xa[idx], ya[idx])
    })
    tibble(n_available = n_available, n_used = n_used,
           entropy_bits = mean(draws), entropy_sd_bits = sd(draws),
           n_draws = ENTROPY_N_DRAWS)
  }) |>
  ungroup() |>
  select(-kid_n_video) |>
  mutate(group = "adults")

entropy_long <- bind_rows(entropy_kids, entropy_adults) |>
  mutate(shot_class = factor(shot_class, levels = SHOTSCALE_LEVELS),
         group = factor(group, levels = c("adults", "kids"))) |>
  select(video_name, group, tile_start_s, tile_end_s, shot_class,
         n_available, n_used, entropy_bits, entropy_sd_bits, n_draws)

cat(glue("{sum(is.na(entropy_long$entropy_bits))} / {nrow(entropy_long)} ",
         "(video, group, tile) cells excluded -- fewer than {ENTROPY_MIN_N} points available\n"))

write_csv(entropy_long, here("data/preprocessed/entropy_by_time.csv"))
cat(glue("Saved -> data/preprocessed/entropy_by_time.csv ({nrow(entropy_long)} rows)\n"))
```

---

# Part 11: Entropy by Shot Scale, by Video

Reads `data/preprocessed/entropy_by_time.csv` directly rather than
recomputing Part 10 -- the entire point of saving that table. Mirrors the
Part 9 ISC-by-shot-scale-by-video layout: one facet per video, raw
(unscaled) means, no cross-video pooling. The one structural difference is
the bars themselves -- two groups (adults, kids) instead of three pairwise
comparisons, since entropy is a within-group statistic.

```{r entropy-plot-load}
entropy_long_plot <- read_csv(here("data/preprocessed/entropy_by_time.csv"), show_col_types = FALSE) |>
  mutate(shot_class = factor(shot_class, levels = SHOTSCALE_LEVELS),
         group = factor(group, levels = c("adults", "kids")))

cat(glue("{sum(is.na(entropy_long_plot$entropy_bits))} / {nrow(entropy_long_plot)} tiles excluded (NA entropy)\n"))

entropy_by_class <- entropy_long_plot |>
  filter(!is.na(entropy_bits)) |>
  group_by(video_name, shot_class, group) |>
  summarise(mean_entropy = mean(entropy_bits, na.rm = TRUE), n_tiles = n(), .groups = "drop")
```

```{r plot-entropy-shotscale, fig.width=18, fig.height=12}
ENTROPY_MAX_BITS <- log2(ENTROPY_GRID_NX * ENTROPY_GRID_NY)

ggplot(entropy_by_class, aes(x = shot_class, y = mean_entropy, fill = group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = ENTROPY_MAX_BITS, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  scale_fill_few(name = "Group") +
  facet_wrap(~video_name, ncol = 4) +
  labs(title = "Fixation-location entropy by shot scale, by video",
       subtitle = glue("4x2 grid, tile-level (0.667s); dashed line = max possible entropy ({round(ENTROPY_MAX_BITS,2)} bits)"),
       x = "Shot scale", y = "Mean entropy (bits)") +
  theme_few(base_size = 12) +
  theme(panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/entropy_by_shotscale.png"), width = 18, height = 12)
```

---

# Part 12: Entropy by Shot Scale, by Block

Same source table and same logic as Part 11, pooled one level up: by block
(Pixar / Frank / Sesame / Slow, via the `video_category()` function defined
in Part 5) instead of by individual video. Reuses `entropy_long_plot` from
Part 11 rather than re-reading the CSV.

```{r entropy-block-aggregate}
entropy_by_block <- entropy_long_plot |>
  filter(!is.na(entropy_bits)) |>
  mutate(block = video_category(video_name)) |>
  group_by(block, shot_class, group) |>
  summarise(mean_entropy = mean(entropy_bits, na.rm = TRUE),
            n_tiles = n(), n_videos = n_distinct(video_name), .groups = "drop")

print(entropy_by_block, n = Inf)
```

```{r plot-entropy-block, fig.width=10, fig.height=8}
ggplot(entropy_by_block, aes(x = shot_class, y = mean_entropy, fill = group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.65, alpha = 0.85) +
  geom_hline(yintercept = ENTROPY_MAX_BITS, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  scale_fill_few(name = "Group") +
  facet_wrap(~block, ncol = 2) +
  labs(title = "Fixation-location entropy by shot scale, by block",
       subtitle = glue("4x2 grid, tile-level (0.667s), pooled across videos within block; ",
                        "dashed line = max possible entropy ({round(ENTROPY_MAX_BITS,2)} bits)"),
       x = "Shot scale", y = "Mean entropy (bits)") +
  theme_few(base_size = 13) +
  theme(panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/entropy_by_shotscale_block.png"), width = 10, height = 8)
```

---

# Part 13: Entropy by Shot Scale, All Videos Averaged

One level further than Part 12: no block or video faceting at all, one pair
of bars per shot-scale class, each bar the unweighted mean across all 12
videos' own per-(shot_class, group) means (same "average of per-video
means, not weighted by time/tiles" convention as every other pooled bar in
this notebook). Individual videos are overlaid as jittered points, coloured
by block, so the spread each bar is averaging over stays visible instead of
disappearing into a single number.

```{r entropy-pooled-aggregate}
bar_stats_entropy <- entropy_by_class |>
  mutate(block = video_category(video_name)) |>
  group_by(shot_class, group) |>
  summarise(mean_entropy = mean(mean_entropy, na.rm = TRUE), n_videos = n(), .groups = "drop")

print(bar_stats_entropy, n = Inf)
```

```{r plot-entropy-pooled, fig.width=11, fig.height=7}
dodge_w <- 0.7
entropy_jit <- position_jitterdodge(jitter.width = 0.12, dodge.width = dodge_w, seed = 42)

ggplot(bar_stats_entropy, aes(x = shot_class, y = mean_entropy, fill = group)) +
  geom_col(position = position_dodge(width = dodge_w), width = 0.65, alpha = 0.7) +
  geom_point(data = entropy_by_class |> mutate(block = video_category(video_name)),
             aes(y = mean_entropy, colour = block, group = group),
             position = entropy_jit, size = 2.2, alpha = 0.85) +
  geom_hline(yintercept = ENTROPY_MAX_BITS, linetype = "dashed", colour = "grey50", linewidth = 0.4) +
  scale_fill_few(name = "Group") +
  scale_colour_few(name = "Block") +
  labs(title = "Fixation-location entropy by shot scale, all videos averaged",
       subtitle = "Bars = unweighted mean of per-video means; points = individual videos by block",
       x = "Shot scale", y = "Mean entropy (bits)") +
  theme_few(base_size = 13) +
  theme(panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/entropy_by_shotscale_pooled.png"), width = 11, height = 7)
```

---

# Part 14: ISC vs. Entropy Timecourse (`r example_video`)

Stacks the windowed ISC timecourse (Part 7/8's 2s-window, 0.25s-step
correlation, reused from `windowed_cache` rather than recomputed) directly
above the tile-level entropy timecourse (0.667s, non-overlapping, from Part
10's saved table), same time axis, to see whether the two measures move
together moment to moment. Entropy is plotted on its own natural scale, not
inverted -- **lower entropy means MORE concentrated/shared gaze**, the
opposite direction from ISC, so a visually concordant pair of signals would
show entropy dipping where ISC peaks, not the two lines tracking each other
directly.

One asymmetry worth keeping in mind when eyeballing the two panels: ISC's
window overlaps 87.5% with its neighbor (2s window, 0.25s step), so it's
smoothed by construction; entropy's tiles are back-to-back with zero
overlap, so it has no such smoothing and will always look jumpier even if
the underlying construct were identical.

```{r isc-entropy-compare-load}
isc_windowed_example <- windowed_cache[[example_video]]

entropy_video <- read_csv(here("data/preprocessed/entropy_by_time.csv"), show_col_types = FALSE) |>
  filter(video_name == example_video, !is.na(entropy_bits)) |>
  mutate(group_label = if_else(group == "adults", "Adult", "Child"))

cat(glue("ISC windows: {nrow(isc_windowed_example)}, entropy tiles: {nrow(entropy_video)}\n"))
```

```{r plot-isc-entropy-compare, fig.width=16, fig.height=8}
p_isc <- ggplot(isc_windowed_example, aes(x = window_start_s, y = mean_r, colour = comparison)) +
  geom_line(linewidth = 0.7) +
  scale_colour_few() +
  labs(y = "Mean ISC (r)", colour = "Comparison", x = NULL,
       title = glue("ISC vs. entropy over time -- {example_video}")) +
  theme_few(base_size = 13) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

p_entropy <- ggplot(entropy_video, aes(x = tile_start_s, y = entropy_bits, colour = group_label)) +
  geom_line(linewidth = 0.7) +
  scale_colour_few() +
  labs(y = "Entropy (bits)", colour = "Group", x = NULL,
       subtitle = "Note: LOWER entropy = MORE concentrated/shared gaze -- opposite direction from ISC above") +
  theme_few(base_size = 13) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

# Adult minus Child entropy, one line -- positive means adults were MORE
# spread out than kids at that moment, negative means kids were more spread
# out. Tile boundaries are identical across groups for a given video by
# construction, so this is a plain join, not another overlap reconciliation.
entropy_diff <- entropy_video |>
  select(tile_start_s, group_label, entropy_bits) |>
  pivot_wider(names_from = group_label, values_from = entropy_bits) |>
  mutate(diff = Adult - Child)

p_diff <- ggplot(entropy_diff, aes(x = tile_start_s, y = diff)) +
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.4) +
  geom_line(linewidth = 0.6, colour = "grey20") +
  labs(y = "Entropy diff\n(Adult - Child, bits)", x = "Time (s)") +
  theme_few(base_size = 13)

p_isc / p_entropy / p_diff + plot_layout(heights = c(1, 1, 0.7))
ggsave(here(glue("figures/isc_vs_entropy_timecourse_{example_video}.png")), width = 16, height = 10)
```

**Quantifying the moment-to-moment agreement**: each ISC window is matched
to whichever entropy tile has the largest time-overlap (same "largest
overlap wins" rule reused throughout this notebook), then correlated within
matched group -- Adult-Adult ISC against Adult entropy, Child-Child ISC
against Child entropy (Child-Adult ISC has no single-group entropy analog,
so it's left out of this comparison).

```{r isc-entropy-correlate}
assign_to_tiles <- function(window_starts, window_s, tile_start_s, tile_end_s) {
  window_ends <- window_starts + window_s
  overlap <- outer(window_ends, tile_end_s, pmin) - outer(window_starts, tile_start_s, pmax)
  overlap[overlap < 0] <- 0
  best <- apply(overlap, 1, which.max)
  best_ov <- apply(overlap, 1, max)
  out <- tile_start_s[best]
  out[best_ov <= 0] <- NA_real_
  out
}

aa_entropy <- entropy_video |> filter(group == "adults") |> select(tile_start_s, entropy_bits)
cc_entropy <- entropy_video |> filter(group == "kids")   |> select(tile_start_s, entropy_bits)

aa_isc <- isc_windowed_example |> filter(comparison == "Adult–Adult") |>
  mutate(tile_start_s = assign_to_tiles(window_start_s, WINDOW_S, aa_entropy$tile_start_s,
                                          aa_entropy$tile_start_s + 0.667)) |>
  inner_join(aa_entropy, by = "tile_start_s")

cc_isc <- isc_windowed_example |> filter(comparison == "Child–Child") |>
  mutate(tile_start_s = assign_to_tiles(window_start_s, WINDOW_S, cc_entropy$tile_start_s,
                                          cc_entropy$tile_start_s + 0.667)) |>
  inner_join(cc_entropy, by = "tile_start_s")

cor_aa <- cor(aa_isc$mean_r, aa_isc$entropy_bits, use = "complete.obs")
cor_cc <- cor(cc_isc$mean_r, cc_isc$entropy_bits, use = "complete.obs")

cat(glue("Adult-Adult ISC vs Adult entropy: r = {round(cor_aa, 3)} (n = {sum(complete.cases(aa_isc))})\n"))
cat(glue("Child-Child ISC vs Child entropy: r = {round(cor_cc, 3)} (n = {sum(complete.cases(cc_isc))})\n"))
cat("(negative r here = concordant: high ISC / low entropy both mean 'more shared gaze')\n")
```

---

# Part 15: Entropy Difference by Shot Scale, All Timepoints Pooled

A different, simpler pooling than Parts 11-13's "unweighted mean of
per-video means": every (video, tile) with a valid entropy value for both
groups gets one Adult-minus-Child diff (same sign convention as Part 14's
diff panel), and the bar is the literal mean over every pooled timepoint --
so a shot-scale class built from more tiles (across all 12 videos combined)
contributes proportionally more, unlike the earlier per-video-then-averaged
bars.

```{r entropy-diff-shotscale-aggregate}
# Tile boundaries are identical across groups for a given video by
# construction, so this is a plain wide-pivot, not another overlap
# reconciliation.
entropy_diff_tiles <- entropy_long_plot |>
  filter(!is.na(entropy_bits)) |>
  select(video_name, tile_start_s, shot_class, group, entropy_bits) |>
  pivot_wider(names_from = group, values_from = entropy_bits) |>
  filter(!is.na(adults), !is.na(kids)) |>
  mutate(diff = adults - kids)

bar_stats_diff <- entropy_diff_tiles |>
  group_by(shot_class) |>
  summarise(mean_diff = mean(diff), se_diff = sd(diff) / sqrt(n()), n_tiles = n(), .groups = "drop")

print(bar_stats_diff, n = Inf)
```

```{r plot-entropy-diff-shotscale, fig.width=9, fig.height=6}
ggplot(bar_stats_diff, aes(x = shot_class, y = mean_diff)) +
  geom_hline(yintercept = 0, colour = "grey60", linewidth = 0.4) +
  geom_col(fill = "#4292c6", width = 0.6, alpha = 0.85) +
  geom_errorbar(aes(ymin = mean_diff - se_diff, ymax = mean_diff + se_diff), width = 0.15, linewidth = 0.7) +
  labs(title = "Entropy difference by shot scale, all timepoints pooled",
       subtitle = "Mean (Adult - Child) entropy per tile, all 12 videos' tiles pooled; error bars = SE",
       x = "Shot scale", y = "Mean entropy diff (Adult - Child, bits)") +
  theme_few(base_size = 13) +
  theme(panel.grid.major.y = element_line(colour = "grey92"))
ggsave(here("figures/entropy_diff_by_shotscale.png"), width = 9, height = 6)
```

---

# Summary

| Step | What we did |
|---|---|
| **QC filter** | Kept participants with ≥ `r min_valid_pct` % valid gaze *and* ≥ `r min_trials` trials (`summary.csv`) |
| **ISC source** | Pre-computed `isc_pairwise.csv`; used `isc_xy = (isc_x + isc_y) / 2` |
| **Time-series** | Loaded raw gaze for `r paste(example_pids, collapse=", ")`; 10 s of `r example_video` |
| **ISCw** | Each observer averaged with all same-age partners per video, then across videos |
| **ISCa** | Each non-adult observer averaged with all adults per video, then across videos |
| **Plots** | Dot plots jittered by subject (Version A) and by video (Version B); bar plot by pair type × video category, Pixar vs. Slow vs. Sesame vs. Frank (Version C); calibration accuracy vs. within-group ISC by category (Part 6); sliding-window ISC timecourse for all three pairings (Part 7, `pixar_birds` + `slow_*`); same timecourse shaded by shot scale instead of high-ISC periods, all 12 videos (Part 8) |
