---
title: "Fixation Density Concordance vs. ISC — Timecourses & Cut-Locked Analysis"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    highlight-style: github
---

# Setup

Reads the three canonical CSVs computed elsewhere (not recomputed here -- this notebook is
plotting only). NOTE: `data/` is a symlink to the mounted lab server in this repo.

- `data/results/concordance/fixation_density_concordance.csv` -- individual vs.
  leave-one-out group, written by `calc_concordance_group.R`.
- `data/results/concordance/fixation_density_concordance_pairwise.csv` -- every pair of
  participants, written by `calc_concordance_pairwise.R`.
- `data/results/concordance/isc_fixation_timecourse.csv` -- windowed
  Adult-Adult/Child-Child/Child-Adult ISC, written by `isc_analysis.qmd`'s "Part 8b" export
  (population-matched to concordance's `valid_data==TRUE` adults -- see
  METHODS_DECISIONS.md, 2026-10-02, for why that match matters and how it was verified).
- `annotations/scene-cuts/all_clips_cuts.csv` -- manually annotated cut timestamps.

Saves figures to `figures/concordance/` (these are real analysis-result figures, not
sanity/parameter-decision checks -- those go in `figures/checks/` instead, e.g. from
`diagnostic_concordance.py`).

"Concordance" here is never "ISC" -- see `calc_concordance_group.R`'s header for the
naming rule (ISC is reserved for the Franchak-style x/y time-series analysis).

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggthemes)
library(glue)
library(here)

CONCORDANCE_GROUP_PATH    <- here("data", "results", "concordance", "fixation_density_concordance.csv")
CONCORDANCE_PAIRWISE_PATH <- here("data", "results", "concordance", "fixation_density_concordance_pairwise.csv")
ISC_PATH                  <- here("data", "results", "concordance", "isc_fixation_timecourse.csv")
CUTS_PATH                 <- here("annotations", "scene-cuts", "all_clips_cuts.csv")
FIGURES_DIR               <- here("figures", "concordance")
dir.create(FIGURES_DIR, showWarnings = FALSE, recursive = TRUE)

# Fixed categorical colors, matching the Python-generated figures from this same analysis
# (figures/concordance/*.png) -- NOT ggthemes::scale_colour_few(), which is reserved
# for the Adult-Adult/Child-Child/Child-Adult comparisons elsewhere in isc_analysis.qmd.
GROUP_COLOR    <- "#2a78d6"  # concordance, vs. leave-one-out group
ISC_COLOR      <- "#eb6834"  # ISC, pairwise
PAIRWISE_COLOR <- "#1baf7a"  # concordance, pairwise

BLOCKS <- list(
  frank  = c("frank_complex", "frank_objects", "frank_play"),
  sesame = c("sesameus_1", "sesameus_2", "sesameindia_1", "sesameindia_2"),
  slow   = c("slow_animals", "slow_hands", "slow_inanimate", "slow_people"),
  pixar  = c("pixar_birds")
)

concordance_group_raw    <- read_csv(CONCORDANCE_GROUP_PATH, show_col_types = FALSE)
concordance_pairwise_raw <- read_csv(CONCORDANCE_PAIRWISE_PATH, show_col_types = FALSE)
isc_raw_tc               <- read_csv(ISC_PATH, show_col_types = FALSE)
cuts_raw                 <- read_csv(CUTS_PATH, show_col_types = FALSE)

isc_theme <- theme_few(base_size = 13) +
  theme(panel.grid.minor.x = element_blank(),
        axis.minor.ticks.x.bottom = element_line(colour = "grey30", linewidth = 0.3),
        axis.minor.ticks.length = unit(3, "pt"))
```

```{r helpers}
# One row per (video, window) for a single metric: concordance-group is averaged across
# participants first (its raw file has one row per participant per window); the other two
# are already one row per window.
get_concordance_group <- function(video) {
  concordance_group_raw |>
    filter(video_name == video) |>
    group_by(window_start_s) |>
    summarise(value = mean(r, na.rm = TRUE), .groups = "drop") |>
    arrange(window_start_s)
}

get_concordance_pairwise <- function(video) {
  concordance_pairwise_raw |>
    filter(video_name == video) |>
    transmute(window_start_s, value = mean_r) |>
    arrange(window_start_s)
}

get_isc <- function(video, comparison = "Adult–Adult") {
  isc_raw_tc |>
    filter(video_name == video, comparison == !!comparison) |>
    transmute(window_start_s, value = mean_r) |>
    arrange(window_start_s)
}

get_cuts <- function(video) {
  sort(cuts_raw$cut_time_s[cuts_raw$video_name == video])
}
```

---

# Part 1: Single-Metric Timecourse

```{r plot-single}
plot_single_timecourse <- function(df, color, title, ylab = "r") {
  ggplot(df, aes(x = window_start_s, y = value)) +
    geom_line(colour = color, linewidth = 0.9) +
    scale_x_continuous(breaks = seq(0, max(df$window_start_s), by = 5),
                        minor_breaks = seq(0, max(df$window_start_s), by = 1),
                        guide = guide_axis(minor.ticks = TRUE)) +
    labs(title = title, x = "Time (s)", y = ylab) +
    isc_theme
}

p1 <- plot_single_timecourse(get_concordance_group("pixar_birds"), GROUP_COLOR,
                        "Concordance (vs. leave-one-out group) — pixar_birds")
p1
ggsave(file.path(FIGURES_DIR, "fixation_density_concordance_timecourse.png"), p1, width = 12, height = 4)
```

---

# Part 2: Concordance vs. ISC Overlay

```{r plot-overlay}
plot_overlay <- function(video, metrics) {
  # metrics: named list of (df, color) pairs, e.g. list(Concordance = list(df=..., colour=...))
  combined <- map_dfr(names(metrics), function(name) {
    metrics[[name]]$df |> mutate(series = name)
  })
  colors <- set_names(map_chr(metrics, "colour"), names(metrics))

  ggplot(combined, aes(x = window_start_s, y = value, colour = series)) +
    geom_line(linewidth = 0.9) +
    scale_colour_manual(values = colors) +
    scale_x_continuous(breaks = seq(0, max(combined$window_start_s), by = 5),
                        minor_breaks = seq(0, max(combined$window_start_s), by = 1),
                        guide = guide_axis(minor.ticks = TRUE)) +
    labs(title = glue("Concordance vs. ISC — {video}"), x = "Time (s)", y = "r", colour = NULL) +
    isc_theme + theme(legend.position = "top")
}

p2 <- plot_overlay("pixar_birds", list(
  "Concordance (group)" = list(df = get_concordance_group("pixar_birds"), colour = GROUP_COLOR),
  "ISC (pairwise)"      = list(df = get_isc("pixar_birds"), colour = ISC_COLOR)
))
p2
ggsave(file.path(FIGURES_DIR, "concordance_vs_isc_timecourse.png"), p2, width = 12, height = 4)
```

---

# Part 3: Three-Way Comparison, with Cuts

```{r plot-three-way}
plot_three_way_with_cuts <- function(video) {
  metrics <- list(
    "ISC (pairwise)"                       = list(df = get_isc(video), colour = ISC_COLOR),
    "Concordance (vs. leave-one-out group)" = list(df = get_concordance_group(video), colour = GROUP_COLOR),
    "Concordance (pairwise)"               = list(df = get_concordance_pairwise(video), colour = PAIRWISE_COLOR)
  )
  cuts <- get_cuts(video)

  plots <- imap(metrics, function(m, name) {
    ggplot(m$df, aes(x = window_start_s, y = value)) +
      geom_vline(xintercept = cuts, colour = "grey60", linewidth = 0.4, linetype = "dashed") +
      geom_line(colour = m$colour, linewidth = 0.9) +
      labs(title = NULL, x = NULL, y = name) +
      isc_theme
  })

  patchwork::wrap_plots(plots, ncol = 1) +
    patchwork::plot_annotation(
      title = glue("ISC / concordance (group) / concordance (pairwise) vs. cuts — {video}"),
      subtitle = glue("{length(cuts)} cuts (dashed lines)")
    )
}

p3 <- plot_three_way_with_cuts("pixar_birds")
p3
ggsave(file.path(FIGURES_DIR, "concordance_isc_vs_cuts_pixar_birds.png"), p3, width = 12, height = 9)
```

---

# Part 4: Cut-Locked Timecourse

Realigns each metric to time-since-most-recent-cut instead of absolute stimulus time,
pooling across every cut (and, for a block, every clip in that block), with a shaded
mean ± SEM band (binning relative time, truncating once too few cuts contribute a bin,
etc.) -- ported from an earlier Python exploration script (since deleted, see
METHODS_DECISIONS.md, 2026-10-02, "No Python plotting scripts for this analysis").

```{r cut-locked-helpers}
# Raw (unbinned) relative-time samples for ONE clip: for each window, time since the most
# recent cut in THIS SAME clip. Windows before the first cut have no reference cut and are
# dropped.
relative_time_samples <- function(df, cuts) {
  if (length(cuts) == 0) return(tibble(rel_time = numeric(0), value = numeric(0)))
  cut_idx <- findInterval(df$window_start_s, cuts)  # 0 = before first cut
  keep <- cut_idx > 0
  tibble(rel_time = df$window_start_s[keep] - cuts[cut_idx[keep]], value = df$value[keep])
}

# Bins (possibly pooled across multiple clips) relative-time samples; truncates to bins
# with at least min_n contributing samples.
bin_stats <- function(samples, bin_s, min_n) {
  samples |>
    mutate(rel_bin = round(rel_time / bin_s) * bin_s) |>
    group_by(rel_bin) |>
    summarise(n = n(), mean = mean(value), sd = sd(value), .groups = "drop") |>
    mutate(sem = sd / sqrt(n)) |>
    filter(n >= min_n) |>
    arrange(rel_bin)
}

# video_timecourses: list of (df, cuts) pairs -- one per clip in scope (a single clip, or
# every clip in a block). Relative time is computed per-clip, pooled before binning.
cut_locked_stats <- function(video_timecourses, bin_s = 0.25, min_n = 5) {
  samples <- map_dfr(video_timecourses, ~ relative_time_samples(.x$df, .x$cuts))
  bin_stats(samples, bin_s, min_n)
}
```

```{r cut-locked-plot-fn}
# scope: either a single video name, or a block name from BLOCKS (pools every clip in it).
build_video_timecourses <- function(scope, metric_fn) {
  videos <- if (scope %in% names(BLOCKS)) BLOCKS[[scope]] else scope
  map(videos, function(v) list(df = metric_fn(v), cuts = get_cuts(v)))
}

plot_cut_locked <- function(scope, bin_s = 0.25, min_n = 5, mode = c("indexed", "raw")) {
  mode <- match.arg(mode)
  metrics <- list(
    "ISC (pairwise)"                       = list(metric_fn = get_isc, colour = ISC_COLOR),
    "Concordance (vs. leave-one-out group)" = list(metric_fn = get_concordance_group, colour = GROUP_COLOR),
    "Concordance (pairwise)"               = list(metric_fn = get_concordance_pairwise, colour = PAIRWISE_COLOR)
  )

  stats_list <- imap(metrics, function(m, name) {
    tcs <- build_video_timecourses(scope, m$metric_fn)
    stats <- cut_locked_stats(tcs, bin_s, min_n)
    t0 <- stats$mean[stats$rel_bin == 0][1]
    if (mode == "indexed") {
      stats <- stats |> mutate(y = mean / t0, y_lo = (mean - sem) / t0, y_hi = (mean + sem) / t0)
    } else {
      stats <- stats |> mutate(y = mean, y_lo = mean - sem, y_hi = mean + sem)
    }
    stats |> mutate(series = name)
  })
  combined <- bind_rows(stats_list)
  colors <- set_names(map_chr(metrics, "colour"), names(metrics))

  n_cuts <- sum(map_int(build_video_timecourses(scope, get_isc), ~ length(.x$cuts)))

  p <- ggplot(combined, aes(x = rel_bin, y = y, colour = series, fill = series)) +
    geom_vline(xintercept = 0, colour = "grey50", linewidth = 0.6, linetype = "dashed") +
    geom_ribbon(aes(ymin = y_lo, ymax = y_hi), alpha = 0.18, colour = NA) +
    geom_line(linewidth = 0.9) +
    scale_colour_manual(values = colors) +
    scale_fill_manual(values = colors) +
    labs(x = "Time since most recent cut (s)",
         y = if (mode == "indexed") "Ratio to value at t=0 (cut onset)" else "r",
         colour = NULL, fill = NULL,
         title = glue("Cut-locked ISC / concordance — {scope}"),
         subtitle = glue("mean ± SEM across {n_cuts} cuts, truncated where fewer than {min_n} cuts contribute",
                          if (mode == "indexed") "; each series indexed to its own t=0 value" else "")) +
    isc_theme + theme(legend.position = "top")

  if (mode == "indexed") p <- p + geom_hline(yintercept = 1, colour = "grey70", linewidth = 0.5, linetype = "dotted")
  p
}

p4_pixar <- plot_cut_locked("pixar_birds", mode = "indexed")
p4_pixar
ggsave(file.path(FIGURES_DIR, "concordance_isc_cut_locked_pixar_birds_combined.png"), p4_pixar, width = 10, height = 5)
```

```{r cut-locked-by-block, fig.height=5}
p4_frank <- plot_cut_locked("frank", mode = "indexed")
p4_frank
ggsave(file.path(FIGURES_DIR, "concordance_isc_cut_locked_frank_combined.png"), p4_frank, width = 10, height = 5)

p4_sesame <- plot_cut_locked("sesame", mode = "indexed")
p4_sesame
ggsave(file.path(FIGURES_DIR, "concordance_isc_cut_locked_sesame_combined.png"), p4_sesame, width = 10, height = 5)

p4_slow <- plot_cut_locked("slow", mode = "indexed")
p4_slow
ggsave(file.path(FIGURES_DIR, "concordance_isc_cut_locked_slow_combined.png"), p4_slow, width = 10, height = 5)
```

```{r cut-locked-raw, fig.height=5}
p4_raw <- plot_cut_locked("pixar_birds", mode = "raw")
p4_raw
ggsave(file.path(FIGURES_DIR, "concordance_isc_cut_locked_pixar_birds_combined_raw.png"), p4_raw, width = 10, height = 5)
```
