---
title: "Inter-Subject Correlation (ISC) — Movie Watching Eye-Tracking"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: cosmo
    highlight-style: github
---

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(glue)
library(patchwork)
library(zoo)
library(here)

# ── Paths ─────────────────────────────────────────────────────────────────────
SERVER_PATH <- "/Volumes/vislearnlab/experiments/movie-watching"
data_dir         <- file.path(SERVER_PATH, "data/raw/")
isc_pairwise_csv <- file.path(SERVER_PATH, "data/data_to_be_analyzed/isc_pairwise.csv")
summary_csv      <- here("data/qc_checks/summary.csv")

# ── QC thresholds 
min_valid_pct <- 50   # mean_validity % across trials
min_trials    <- 6    # minimum usable trials (half of total trials)

# ── Example time-series participants 
# Two adults and one child used for the 10-second gaze trace plot.
# Change these to any valid participant IDs in your dataset.
ex_adult_1 <- "MW047"
ex_adult_2 <- "MW048"
ex_child   <- "MWK006"

example_video <- "sesameus_2"   # video shown in the time-series panel
```

---

# Part 1: QC Filtering

We use the pre-computed `summary.csv` to identify which participants have
sufficient data quality, then filter `isc_pairwise.csv` to only those pairs
where **both** participants pass QC.

```{r qc}
summ <- read_csv(summary_csv, show_col_types = FALSE) |>
  select(participant_id, n_trials, mean_validity)

passed_qc <- summ |>
  filter(mean_validity >= min_valid_pct, n_trials >= min_trials) |>
  filter(!startsWith(participant_id, "HMET")) |> #excluding infant participants for now
  pull(participant_id)

cat(glue("{length(passed_qc)} / {nrow(summ)} participants pass QC ",
         "(≥{min_valid_pct}% valid, ≥{min_trials} trials)\n\n"))

summ |>
  mutate(included = participant_id %in% passed_qc) |>
  arrange(desc(included), desc(mean_validity))
```

---

# Part 2: Load & Filter Pairwise ISC

```{r load-isc}
isc_raw <- read_csv(isc_pairwise_csv, show_col_types = FALSE)

# Keep only pairs where both participants passed QC
isc <- isc_raw |>
  filter(pid_a %in% passed_qc, pid_b %in% passed_qc) |>
  # Use isc_xy (average of horizontal + vertical) as the primary ISC measure
  rename(isc = isc_xy) |>
  # Simplify comparison labels for plotting
  mutate(
    pair_type = case_when(
      comparison == "adults-adults"   ~ "Adult–Adult",
      comparison == "kids-kids"       ~ "Child–Child",
      comparison == "adults-kids"     ~ "Child–Adult",
      comparison == "infants-infants" ~ "Infant–Infant",
      comparison == "adults-infants"  ~ "Adult–Infant",
      comparison == "infants-kids"    ~ "Infant–Child",
      TRUE                             ~ comparison
    ),
    pair_type = factor(pair_type, levels = c(
      "Infant–Infant", "Infant–Child", "Child–Child",
      "Child–Adult", "Adult–Infant", "Adult–Adult"
    ))
  )

cat(glue("Retained {nrow(isc)} pair × video rows ",
         "({n_distinct(c(isc$pid_a, isc$pid_b))} unique participants)\n"))
isc |> count(pair_type)
```

---

# Part 3: Example Time-Series (`r example_video`, first 10 s)

We load the raw gaze files for the three example participants directly.

```{r ts-load-helpers}
# ── File discovery: participant ID is the prefix of the filename ──────────────
find_gaze_path <- function(pid) {
  hits <- list.files(data_dir, pattern = glue("^{pid}_.*\\.csv$"),
                     full.names = TRUE, recursive = TRUE) |>
    keep(~ !str_detect(basename(.x), "_trial_order|_validation"))
  if (length(hits) == 0) { warning(glue("No gaze file found for {pid}")); return(NULL) }
  hits[which.min(nchar(basename(hits)))]
}

find_trial_order_path <- function(pid) {
  hits <- list.files(data_dir, pattern = glue("^{pid}_.*_trial_order\\.csv$"),
                     full.names = TRUE, recursive = TRUE)
  if (length(hits) == 0) { warning(glue("No trial_order found for {pid}")); return(NULL) }
  hits[1]
}

load_gaze_pid <- function(pid) {
  path <- find_gaze_path(pid)
  if (is.null(path)) return(NULL)
  read_csv(path, show_col_types = FALSE) |> mutate(participant = pid)
}

# ── Load the three example participants ───────────────────────────────────────
example_pids <- c(ex_adult_1, ex_adult_2, ex_child)

gaze_ex <- map(example_pids, load_gaze_pid) |>
  compact() |>
  list_rbind()

# Use one participant's trial order for the video -> trial_index lookup
to_path <- find_trial_order_path(ex_adult_1)
trial_order_ex <- read_csv(to_path, show_col_types = FALSE)

video_key <- trial_order_ex |>
  select(trial_index = total_trial_index, video_name, block_id)

cat("Loaded gaze for:", paste(unique(gaze_ex$participant), collapse = ", "), "\n")
```

```{r ts-parse}
# Label rows with trial / video (same logic as before)
gaze_labeled_ex <- gaze_ex |>
  group_by(participant) |>
  mutate(
    video_name = str_extract(events, "(?<=\\|Video_)[^|]+"),
    video_name = zoo::na.locf(video_name, na.rm = FALSE)
  ) |>
  ungroup() |>
  left_join(video_key, by = "video_name") |>
  filter(!is.na(trial_time), !is.na(trial_index))

# Resample to 8 ms grid (4 ms == 250 Hz)
gaze_resampled_ex <- gaze_labeled_ex |>
  mutate(time_bin = round(trial_time / 8) * 8) |>
  group_by(participant, trial_index, time_bin) |>
  summarise(gaze_x = mean(gaze_x, na.rm = TRUE),
            gaze_y = mean(gaze_y, na.rm = TRUE),
            .groups = "drop")
```

```{r ts-plot-helper}
plot_isc_timeseries <- function(p1, p2, label1, label2, panel_title,
                                video = example_video, max_sec = 10) {
  t_idx <- video_key |> filter(video_name == video) |> pull(trial_index)
  if (length(t_idx) == 0) {
    message("Video '", video, "' not found. Available: ",
            paste(sort(unique(video_key$video_name)), collapse = ", "))
    return(NULL)
  }
  t_idx <- t_idx[1]

  ts_data <- gaze_resampled_ex |>
    filter(participant %in% c(p1, p2),
           trial_index == t_idx,
           time_bin <= max_sec * 1000) |>
    group_by(participant) |>
    mutate(gaze_x_norm = (gaze_x - min(gaze_x, na.rm = TRUE)) /
             (max(gaze_x, na.rm = TRUE) - min(gaze_x, na.rm = TRUE))) |>
    ungroup() |>
    mutate(participant = factor(participant, levels = c(p1, p2)))

  # Pull ISC value from the pre-computed pairwise table
  isc_val <- isc |>
    filter(
      (pid_a == p1 & pid_b == p2) | (pid_a == p2 & pid_b == p1),
      video_name == video
    ) |>
    pull(isc_x) |> first() |> round(2)
  isc_label <- if (is.na(isc_val)) "ISC = NA" else glue("horizontal ISC = {isc_val}")

  ggplot(ts_data, aes(x = time_bin / 1000, y = gaze_x_norm,
                      colour = participant, linewidth = participant)) +
    geom_line(na.rm = TRUE) +
    scale_colour_manual(
      values = setNames(c("grey60", "black"), c(p1, p2)),
      labels = setNames(c(label1, label2), c(p1, p2))
    ) +
    scale_linewidth_manual(
      values = setNames(c(0.7, 1.2), c(p1, p2)),
      labels = setNames(c(label1, label2), c(p1, p2))
    ) +
    annotate("text", x = max_sec * 0.97, y = 0.05,
             label = isc_label, hjust = 1, size = 3.5) +
    scale_x_continuous(breaks = seq(0, max_sec, 2)) +
    scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.05, 1.05)) +
    labs(title = panel_title, x = "Time (s)",
         y = "Horizontal gaze position",
         colour = NULL, linewidth = NULL) +
    theme_classic(base_size = 12) +
    theme(legend.position = c(0.08, 0.25))
}
```

```{r ts-plots, fig.width=8, fig.height=7, warning=FALSE}
p_aa <- plot_isc_timeseries(
  ex_adult_1, ex_adult_2, "A1", "A2",
  panel_title = glue("(a) Adult–Adult ISC ({example_video})")
)
p_ca <- plot_isc_timeseries(
  ex_adult_1, ex_child, "A1", "C1",
  panel_title = glue("(b) Child–Adult ISC ({example_video})")
)
p_aa / p_ca
ggsave(here("figures/isc_correlation.png"))
```

::: {.callout-tip}
**What to notice**

- In the adult–adult panel the traces closely follow each other — both adults tend to look at the same screen region at the same time.
- In the child–adult panel the traces diverge more, reflected in a lower ISC value.
:::

---

# Part 4: ISCw and ISCa

Following Franchak et al. (2016):

- **ISCw** — each observer's mean ISC with all *same-age* partners, averaged across videos.
- **ISCa** — each child's mean ISC with all *adult* partners, averaged across videos.

```{r isc-averages}
# Reshape to one row per focal observer × partner × video
isc_long <- isc |>
  mutate(
    partner_a = pid_b,
    partner_b = pid_a
  ) |>
  pivot_longer(
    cols = c(pid_a, pid_b),
    names_to = "role",
    values_to = "focal"
  ) |>
  mutate(
    partner = if_else(role == "pid_a", partner_a, partner_b),
    focal_group = if_else(role == "pid_a", group_a, group_b),
    partner_group = if_else(role == "pid_a", group_b, group_a)
  ) |>
    select(
    focal,
    partner,
    video_name,
    isc,
    pair_type,
    focal_group,
    partner_group
  )

# ISCw: within-age, averaged over partners then over videos
iscw_by_video <- isc_long |>
  filter(focal_group == partner_group) |>
  group_by(focal, focal_group, video_name) |>
  summarise(ISCw = mean(isc, na.rm = TRUE), .groups = "drop")

iscw_by_subj <- iscw_by_video |>
  group_by(focal, focal_group) |>
  summarise(ISCw = mean(ISCw, na.rm = TRUE), .groups = "drop")

# ISCa: each non-adult observer averaged with all adults
isca_by_video <- isc_long |>
  filter(focal_group != "adults", partner_group == "adults") |>
  group_by(focal, focal_group, video_name) |>
  summarise(ISCa = mean(isc, na.rm = TRUE), .groups = "drop")

isca_by_subj <- isca_by_video |>
  group_by(focal, focal_group) |>
  summarise(ISCa = mean(ISCa, na.rm = TRUE), .groups = "drop")
```

---

# Part 5: Summary Plots

Two versions of a Franchak et al.–style dot plot. Mean ± 1 SE shown as a
filled point + error bars; grey dots are individual observations.

**Version A** — one dot per *subject* (ISC averaged over videos per subject)  
**Version B** — one dot per *video* (ISC averaged over subjects per video)

```{r build-plot-data}
# We plot only the three pairings that mirror Franchak Fig. 2:
# kids-kids, adults-kids, adults-adults
# (infants shown separately if present)

plot_comparisons <- c("Child–Child", "Child–Adult", "Adult–Adult")

# ── Version A: per subject ────────────────────────────────────────────────────
subj_avg <- bind_rows(
  # Adult–Adult: each adult's mean ISC with all other adults
  isc_long |>
    filter(focal_group == "adults", partner_group == "adults") |>
    group_by(focal) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Adult–Adult"),

  # Child–Child
  isc_long |>
    filter(focal_group == "kids", partner_group == "kids") |>
    group_by(focal) |>
    summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
    mutate(pair_type = "Child–Child"),

  # Child–Adult (from child's perspective)
  isca_by_subj |>
    filter(focal_group == "kids") |>
    rename(isc_mean = ISCa) |>
    select(focal, isc_mean) |>
    mutate(pair_type = "Child–Adult")
) |>
  mutate(pair_type = factor(pair_type, levels = plot_comparisons))

# ── Version B: per video ──────────────────────────────────────────────────────
video_avg <- isc |>
  filter(pair_type %in% plot_comparisons) |>
  group_by(pair_type, video_name) |>
  summarise(isc_mean = mean(isc, na.rm = TRUE), .groups = "drop") |>
  mutate(pair_type = factor(pair_type, levels = plot_comparisons))

# ── Summary stats ─────────────────────────────────────────────────────────────
se_stats <- function(df) {
  df |>
    group_by(pair_type) |>
    summarise(
      mean_isc = mean(isc_mean, na.rm = TRUE),
      se_isc   = sd(isc_mean, na.rm = TRUE) / sqrt(sum(!is.na(isc_mean))),
      .groups  = "drop"
    )
}

subj_stats  <- se_stats(subj_avg)
video_stats <- se_stats(video_avg)
```

```{r plot-theme}
isc_theme <- theme_classic(base_size = 13) +
  theme(
    axis.text.x        = element_text(size = 11),
    legend.position    = "none",
    panel.grid.major.y = element_line(colour = "grey92")
  )

dot_plot <- function(stats_df, jitter_df, plot_title) {
  ggplot(stats_df, aes(x = pair_type, y = mean_isc)) +
    geom_jitter(data  = jitter_df, aes(y = isc_mean),
                width = 0.12, size = 1.8, alpha = 0.5, colour = "grey55") +
    geom_errorbar(aes(ymin = mean_isc - se_isc, ymax = mean_isc + se_isc),
                  width = 0.12, linewidth = 0.8) +
    geom_point(size = 4) +
    scale_y_continuous(breaks = seq(0, 1, 0.1)) +
    labs(title = plot_title, x = NULL, y = "ISC (Pearson r)") +
    isc_theme
}
```

## 5.1 Version A — jitter by subject

```{r plot-subj, fig.width=5, fig.height=5}
dot_plot(subj_stats, subj_avg, "Inter-subject correlation\n(dots = subjects)")
ggsave(here("figures/isc_subjects.png"))
```

## 5.2 Version B — jitter by video

```{r plot-video, fig.width=10, fig.height=10, dpi=300}
  ggplot(video_stats, aes(x = pair_type, y = mean_isc)) +
    geom_jitter(data  = video_avg, aes(y = isc_mean, color=video_name),
                width = 0.12, size = 4, alpha = 0.5) +
    geom_errorbar(aes(ymin = mean_isc - se_isc, ymax = mean_isc + se_isc),
                  width = 0.12, linewidth = 0.8) +
    geom_point(size = 4) +
    #scale_color_viridis_d() +
    scale_y_continuous(breaks = seq(0, 1, 0.1)) +
    labs(title = "ISC by pair type\n(dots = videos)", x = NULL, y = "ISC (Pearson r)")
ggsave(here("figures/isc_videos.png"))
```

---

# Summary

| Step | What we did |
|---|---|
| **QC filter** | Kept participants with ≥ `r min_valid_pct` % valid gaze *and* ≥ `r min_trials` trials (`summary.csv`) |
| **ISC source** | Pre-computed `isc_pairwise.csv`; used `isc_xy = (isc_x + isc_y) / 2` |
| **Time-series** | Loaded raw gaze for `r paste(example_pids, collapse=", ")`; 10 s of `r example_video` |
| **ISCw** | Each observer averaged with all same-age partners per video, then across videos |
| **ISCa** | Each non-adult observer averaged with all adults per video, then across videos |
| **Plots** | Two dot-plot versions: jitter by subject (Version A) and by video (Version B) |
